---
title: "QC ChromBPNet models"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: 
    code_folding: show
    theme: flatly
    highlight: pygments
    toc: yes
    toc_float:
      collapsed: true
    toc_depth: 4
    number_sections: yes
    keep_md: no
---

```{r config, warning = FALSE, message = FALSE}

# **MODIFY THIS CHUNK**
library(here)
proj_dir    <- here()
doc_id      <- "03"
out         <- here("output/03-chrombpnet/01-models/qc"); dir.create(out, recursive = TRUE)
figout      <- here("figures/03-chrombpnet/01-models", doc_id, "/"); dir.create(figout, recursive = TRUE)

```

```{r setup, include = FALSE}

# NO NEED TO MODIFY THIS CHUNK

knitr::knit_hooks$set( echo_label = function(before, options, envir) {
  if ( before ) {
    # Do nothing
  } else sprintf('<br><span style="color:#1abc9c">~[output @ *%s*]~</span>',
                 paste0(options$fig.path, "/", options$label, "...") )
})

knitr::opts_chunk$set(message = TRUE,
                      warning = FALSE,
                      error = FALSE,
                      cache = FALSE,
                      fig.path = figout,
                      fig.keep = "all",
                      dev = c("png", "pdf"),
                      cache.lazy = FALSE)

grDevices::pdf.options(useDingbats = FALSE)

options(knitr.table.format = "html") 
knitr::opts_knit$set(width = 1600)

set.seed(100)

```

***

# Overview

Here we load and aggregate the QC metrics for the ChromBPNet models to evaluate them prior
to interpretation. We set thresholds to define the set of models which will
be used for downstream analysis.


# Set up

```{r libs, warning = FALSE, message = FALSE, echo = TRUE, cache = FALSE}

library(dplyr)
library(tidyr)
library(ggplot2)
library(readr)
library(scales)
library(glue)
library(purrr)
library(stringr)
library(cowplot)

script_path <- here("code/utils/")
source(file.path(script_path, "plotting_config.R"))
source(file.path(script_path, "hdma_palettes.R"))
source(file.path(script_path, "sj_scRNAseq_helpers.R"))

ggplot2::theme_set(theme_min())

```


# Load QC and visualize

Load general cluster info (number of fragments, number of cells):

```{r}

# get number of cells
cluster_meta <- read_csv(here("output/05-misc/03/TableS2_cluster_meta_qc.csv"))

# get number of fragments
n_frags <- read_tsv(here("output/01-preprocessing/03/fragmentsPerCluster.tsv")) %>% 
  dplyr::rename(total_frags = total_reads)

# put these together
cluster_meta <- cluster_meta %>% 
  left_join(n_frags, by = c("Cluster" = "RNA_Clusters")) %>% 
  # in the ChromBPNet analyses, StomachEsophagus is referred to as Stomach
  mutate(organ = ifelse(organ == "StomachEsophagus", "Stomach", organ))

sorted_cluster_order <- cluster_meta %>% arrange(organ, cluster_id) %>% pull(Cluster_ChromBPNet)

```


Then, we load the various ChromBPNet QC metrics:

```{r load}

bias_model <- "Heart_c0_thresh0.4"
model_dir  <- here("output/03-chrombpnet/01-models/models/", paste0("bias_", bias_model))
fs::dir_exists(model_dir)

metrics    <- fs::dir_ls(model_dir, glob = "*/*/evaluation/chrombpnet_metrics.json", recurse = TRUE, type = "file")

# extract out cell type and fold from the path
clusters   <- stringr::str_split(metrics, "/") %>% sapply(getElement, 14)
folds      <- stringr::str_split(metrics, "/") %>% sapply(getElement, 15)

# load the response of the bias model
bias_resp  <- fs::dir_ls(model_dir, glob = "*/*/evaluation/chrombpnet_nobias_max_bias_response.txt",
                         recurse = TRUE, type = "file")

# loop over model performance metrics and read them in
chrombpnet_metrics <- pmap_dfr(list(metrics, clusters, folds),
                               ~ jsonlite::fromJSON(..1) %>%
                                 as.data.frame() %>% 
                                 tibble::add_column("Cluster" = ..2, .before = 1) %>% 
                                 tibble::add_column("Fold" = ..3, .after = 1))

# get tn5 motif response
read_bias <- function(response_file) {
  
  responses <- read_lines(response_file) %>% str_split("_") %>% .[[1]] %>% getElement(3) %>% str_split("/") %>% .[[1]]
  data.frame("tn5_motif" = paste0("tn5_motif_", 1:5),
             "response" = responses) %>% 
    pivot_wider(names_from = "tn5_motif", values_from = "response") %>% 
    mutate(Cluster = response_file %>% str_split("/") %>% .[[1]] %>% getElement(14),
           Fold = response_file %>% str_split("/") %>% .[[1]] %>% getElement(15))
  
}

bias_metrics <- map_dfr(bias_resp, read_bias)

```


Put all metrics together:

```{r}

all_metrics <- right_join(cluster_meta, chrombpnet_metrics, by = c("Cluster_ChromBPNet" = "Cluster")) %>% 
  left_join(bias_metrics, by = c("Cluster_ChromBPNet" = "Cluster", "Fold" = "Fold")) %>% 
  mutate(Model = paste0(Cluster_ChromBPNet, "_", Fold)) %>% 
  dplyr::relocate(Model, .before = 1) %>% 
  arrange(organ, L1_annot) %>% 
  rowwise() %>% 
  mutate(max_tn5_response = max(tn5_motif_1, tn5_motif_2, tn5_motif_3, tn5_motif_4, tn5_motif_5))

head(all_metrics)

```


Sanity checks:

```{r}

n_clusters <- 203
length(unique(all_metrics$Model)) == 5 * n_clusters
length(unique(all_metrics$Cluster)) == n_clusters
length(unique(all_metrics$organ)) == 12

```

Filter out cell types where any of the folds has < 0.5 Spearman counts correlation.

```{r}

celltypes_drop <- all_metrics %>%
  filter(counts_metrics.peaks.spearmanr < 0.5) %>% 
  pull(Cluster) %>% 
  unique()
celltypes_drop

all_metrics <- all_metrics %>% 
  mutate(Keep_cluster = ifelse(Cluster %in% celltypes_drop, FALSE, TRUE),
         Keep_fold    = case_when(
           Cluster %in% celltypes_drop ~ FALSE,
           TRUE ~ TRUE
         ))

models_to_keep <- all_metrics %>%
  dplyr::select(Model, Cluster_ChromBPNet, Cluster, Fold, Keep_cluster, Keep_fold) %>%
  group_by(Cluster_ChromBPNet, Cluster) %>%
  mutate(Keep_cluster = ifelse(all(Keep_cluster), TRUE, FALSE)) %>%
  filter(Keep_fold) %>%
  summarize(Folds_to_keep = glue_collapse(Fold, ",")) %>% 
  relocate(Cluster, .after = Folds_to_keep)

models_to_keep %>% write_tsv(glue("{out}/chrombpnet_models_keep.tsv"), col_names = FALSE)

nrow(models_to_keep)

```

Thus, at this stage, we're keeping 199 models.

Prepare supplementary table:

```{r table}

all_metrics %>% 
  dplyr::select(organ, Cluster_ChromBPNet, L1_annot, number_of_cells = ncell, total_frags, Model, Fold, matches("metrics"), matches("tn5")) %>% 
  write_tsv(glue("{out}/chrombpnet_metrics.tsv"))

```


Correlation on the counts task:

```{r}

summary(all_metrics$counts_metrics.peaks.pearsonr)
summary(all_metrics$counts_metrics.peaks.spearmanr)
summary(all_metrics$profile_metrics.peaks.median_jsd)

```



```{r chrombpnet_qc_metrics_barplots, fig.width = 30, fig.height = 15}

cluster_meta <- cluster_meta %>% 
  mutate(Cluster_ChromBPNet = factor(Cluster_ChromBPNet, levels = sorted_cluster_order))

all_metrics_avg <- all_metrics %>% 
  group_by(Cluster_ChromBPNet, Cluster, dend_order, organ) %>% 
  summarize_at(.vars = vars(counts_metrics.peaks.pearsonr, counts_metrics.peaks.spearmanr, profile_metrics.peaks.median_jsd), .funs = mean) %>% 
  gather(stat, value, counts_metrics.peaks.pearsonr, counts_metrics.peaks.spearmanr, profile_metrics.peaks.median_jsd)

all_metrics_long <- all_metrics %>%
  gather(stat, value, counts_metrics.peaks.pearsonr, counts_metrics.peaks.spearmanr, profile_metrics.peaks.median_jsd)

p1 <- cluster_meta %>% 
  ggplot(aes(x = Cluster_ChromBPNet, y = total_frags)) +
  geom_bar(aes(fill = organ), stat = "identity") +
  geom_hline(yintercept = c(5e6), color = "black", linetype = "dashed") +
  scale_fill_manual(values = cmap_organ) +
  # put the scale on a log10 scale, but use non-logged numbers
  scale_y_log10(label = comma) +
  coord_cartesian(ylim = c(1e5, NA)) +
  facet_grid(. ~ organ, scales = "free_x", space = "free") +
  rotate_x() + no_legend() +
  theme(axis.text.x = element_blank(),
        panel.grid.major.y = element_line(color = "gray90")) +
  xlab(NULL) + ylab("Total # fragments") +
  ggtitle("ChromBPNet model performance")

p2 <- cluster_meta %>% 
  ggplot(aes(x = Cluster_ChromBPNet, y = ncell)) +
  geom_bar(aes(fill = organ), stat = "identity") +
  scale_fill_manual(values = cmap_organ) +
  # put the scale on a log10 scale, but use non-logged numbers
  scale_y_log10(label = comma) +
  facet_grid(. ~ organ, scales = "free_x", space = "free") +
  rotate_x() + no_legend() +
  theme(axis.text.x = element_blank(),
        strip.text.x = element_blank(),
        panel.grid.major.y = element_line(color = "gray90")) +
  xlab(NULL) + ylab("# cells")

p3 <- all_metrics_avg %>% 
  mutate(Cluster_ChromBPNet = factor(Cluster_ChromBPNet, levels = sorted_cluster_order)) %>% 
  filter(stat %in% c("counts_metrics.peaks.pearsonr")) %>% 
  mutate(stat = recode(stat, counts_metrics.peaks.spearmanr = "SpearmanR", counts_metrics.peaks.pearsonr = "PearsonR")) %>% 
  ggplot(aes(x = Cluster_ChromBPNet, y = value)) +
  geom_bar(aes(fill = organ), stat = "identity") +
  geom_point(data = all_metrics_long %>%
               filter(stat %in% c("counts_metrics.peaks.pearsonr")) %>% 
               mutate(stat = recode(stat, counts_metrics.peaks.spearmanr = "SpearmanR", counts_metrics.peaks.pearsonr = "PearsonR")), 
             aes(x = Cluster_ChromBPNet, y = value), shape = 21, size = 0.5) +
  scale_fill_manual(values = cmap_organ) +
  # geom_text(aes(label = round(value, 2), y = 0.88), size = 2, angle = 90) +
  facet_grid(. ~ organ, scales = "free_x", space = "free") +
  rotate_x() +
  scale_y_continuous(breaks = seq(0.3, 0.9, 0.1)) +
  coord_cartesian(ylim = c(0.3, 0.9)) +
  ylab("Correlation between predicted / observed counts") + no_legend() +
  theme(strip.text.x = element_blank(),
        panel.grid.major.y = element_line(color = "gray90"))

p4 <- all_metrics_avg %>% 
  mutate(Cluster_ChromBPNet = factor(Cluster_ChromBPNet, levels = sorted_cluster_order)) %>% 
  filter(stat %in% c("profile_metrics.peaks.median_jsd")) %>%
  mutate(stat = "Median JSD") %>% 
  ggplot(aes(x = Cluster_ChromBPNet, y = value)) +
  geom_bar(aes(fill = organ), stat = "identity") +
  geom_point(data = all_metrics_long %>%
               filter(stat %in% c("profile_metrics.peaks.median_jsd")) %>% 
               mutate(stat = "Median JSD"),
             aes(x = Cluster_ChromBPNet, y = value), shape = 21, size = 0.5) +
  scale_fill_manual(values = cmap_organ) +
  facet_grid(stat ~ organ, scales = "free_x", space = "free") +
  rotate_x() +
  coord_cartesian(ylim = c(0.3, 0.85)) +
  scale_y_continuous(breaks = seq(0.3, 0.85, 0.1)) +
  ylab("Median JSD between observed / predicted profiles") + no_legend() +
  theme(strip.text.x = element_blank(),
        panel.grid.major.y = element_line(color = "gray90")) +
  xlab("Cluster")

plot_grid(p1, p2, p3 + theme(axis.text.x = element_blank()) + xlab(NULL),
          p4, ncol = 1, rel_heights = c(1.5, 1, 1, 2), align = "v", axis = "rl")

```


Plot response to Tn5 motifs:

```{r model_tn5_response, fig.width = 30, fig.height = 11}

all_metrics %>% 
  mutate(Cluster_ChromBPNet = factor(Cluster_ChromBPNet, levels = sorted_cluster_order)) %>% 
  dplyr::select(Cluster_ChromBPNet, organ, Fold, matches("tn5_motif")) %>% 
  gather(Tn5_motif, Response, matches("tn5_motif")) %>% 
  mutate(Response = as.numeric(Response),
         Response = ifelse(Response > 0.01, 0.01, Response)) %>% 
  ggplot(aes(x = Cluster_ChromBPNet, y = Response)) +
  geom_hline(yintercept = 0.003, color = "gray90") +
  geom_jitter(stat = "identity", aes(color = organ, shape = Fold), width = 0.1, alpha = 0.8) +
  scale_y_continuous(breaks = seq(0.001, 0.01, 0.002)) +
  facet_grid(Tn5_motif ~ organ, scales = "free_x", space = "free") +
  scale_color_manual(values = cmap_organ) + 
  rotate_x() +
  ggtitle("ChromBPNet Tn5 motif response, plotting capped at 0.01") +
  coord_cartesian(ylim = c(0, 0.01))

```

Distribution of metrics (mean across folds) per cell type:

```{r metrics_distributions, fig.width = 9, fig.height = 6}

p1 <- all_metrics_avg %>%
  filter(stat == "counts_metrics.peaks.pearsonr") %>%
  ggplot(aes(x = 1, y = value)) +
  geom_boxplot(outlier.colour = NA, fill = "gray90") +
  geom_jitter(fill = "#3953A4", width = 0.25,
              shape = 21, size = 3, alpha = 0.9) +
  theme(
    panel.grid.major.y = element_line(),
    panel.grid.minor.y = element_line(),
    axis.ticks.x = element_blank(),
    axis.text.x = element_blank()) +
  ggtitle("Pearson R") + ylab("Value") + xlab(NULL)

p2 <- all_metrics_avg %>%
  filter(stat == "counts_metrics.peaks.spearmanr") %>%
  ggplot(aes(x = 1, y = value)) +
  geom_boxplot(outlier.colour = NA, fill = "gray90") +
  geom_jitter(fill = "#3953A4", width = 0.25,
              shape = 21, size = 3, alpha = 0.9) +
  theme(
    panel.grid.major.y = element_line(),
    panel.grid.minor.y = element_line(),
    axis.ticks.x = element_blank(),
    axis.text.x = element_blank()) +
  ggtitle("Spearman R") + ylab("Value") + xlab(NULL)

p3 <- all_metrics_avg %>%
  filter(stat == "profile_metrics.peaks.median_jsd") %>%
  ggplot(aes(x = 1, y = value)) +
  geom_boxplot(outlier.colour = NA, fill = "gray90") +
  geom_jitter(fill = "#3953A4", width = 0.25,
              shape = 21, size = 3, alpha = 0.9) +
  theme(
    panel.grid.major.y = element_line(),
    panel.grid.minor.y = element_line(),
    axis.ticks.x = element_blank(),
    axis.text.x = element_blank()) +
  ggtitle("Median JSD") + ylab("Value") + xlab(NULL)

plot_grid(p1, p2, p3, nrow = 1)

```



# Session info

```{r sinfo, cache = FALSE}

.libPaths()
sessionInfo()

```

