---
title: "Global per-sample QC"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: 
    code_folding: show
    theme: flatly
    highlight: pygments
    toc: yes
    toc_float:
      collapsed: true
    toc_depth: 4
    number_sections: yes
    keep_md: no
---

```{r config, warning = FALSE}

# **MODIFY THIS CHUNK**
project_id  <- "HDMA-public" # determines the name of the cache folder
doc_id      <- "02-global_analysis/01" # determines name of the subfolder of `figures` where pngs/pdfs are saved
out         <- here::here("output/", doc_id); dir.create(out, recursive = TRUE)
figout      <- here::here("figures/", doc_id, "/")

```

```{r setup, include = FALSE}

# NO NEED TO MODIFY THIS CHUNK

knitr::knit_hooks$set( echo_label = function(before, options, envir) {
  if ( before ) {
    # Do nothing
  } else sprintf('<br><span style="color:#1abc9c">~[output @ *%s*]~</span>',
                 paste0(options$fig.path, "/", options$label, "...") )
})

knitr::opts_chunk$set(message = TRUE,
                      warning = FALSE,
                      error = FALSE,
                      cache = FALSE,
                      fig.path = figout,
                      fig.keep = "all",
                      dev = c("png", "pdf"),
                      cache.lazy = FALSE)

grDevices::pdf.options(useDingbats = FALSE)

options(knitr.table.format = "html") 
knitr::opts_knit$set(width = 1600)

set.seed(100)

```

***

# Overview

Global overview of the HDMA atlas, including QC stats and metadata at the
tissue and sample levels.


# Set up

```{r libs, warning = FALSE, message = FALSE, echo = TRUE, cache = FALSE}

# libraries
library(glue)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(purrr)
library(readr)
library(tibble)
library(stringr)
library(Seurat)
library(cowplot)
library(data.table)
library(ArchR)
library(here)

# shared project scripts/functions
source(here::here("code/utils/plotting_config.R"))
source(here::here("code/utils/matrix_helpers.R"))
source(here::here("code/utils/seurat_helpers.R"))
source(here::here("code/utils/sj_scRNAseq_helpers.R"))
source(here::here("code/utils/hdma_palettes.R"))

ggplot2::theme_set(theme_BOR())

```


# Extract metadata

```{r}

# TSV file to manually define which organs to keep
organs_keep <- read_tsv(here::here("code/02-global_analysis/01-organs_keep.tsv"))
organs_keep

```

Set some paths:

```{r}

preprocessing_out_dir <- here::here("output/01-preprocessing/02")
meta_dir <- file.path(out, "per_cell_meta")
dir.create(meta_dir, showWarnings = F, recursive = T)

```


Extract RNA metadata:

```{r rna_meta}

organs <- organs_keep$organ
organ_codes <- organs_keep$organcode

seurat_filepaths <- map(organs, ~ file.path(preprocessing_out_dir, .x, "rna_preprocess_output/RNA_obj_clustered_final.rds")) %>% 
  set_names(organs)

# check all files exist
all(map_lgl(seurat_filepaths, file.exists))

# create a TSV file for metadata from each organ 
for (organ in organs) {
  
  filepath <- seurat_filepaths[[organ]]
  # print(filepath)
  
  rna_meta_path <- file.path(meta_dir, paste0("/", organ, "_rna_meta.tsv"))
  # print(rna_meta_path)
  
  # only process if the Seurat object exists, and the meta doesn't
  if (file.exists(filepath) & !file.exists(rna_meta_path)) {
    
    message(paste0("Loading Seurat object for ", organ, "..."))
    so <- readRDS(filepath)
    so_meta <- so@meta.data
    write_tsv(so_meta, rna_meta_path)
    
  } else message(organ, " meta already exists, skipping...")
  
}

```


Extract ATAC metadata

```{r atac_meta}

archr_filepaths <- map(organs, ~ file.path(preprocessing_out_dir, .x, "atac_preprocess_output/ATAC_obj_clustered_peaks_final")) %>% 
  set_names(organs)

# check all files exist
all(map_lgl(archr_filepaths, file.exists))

# Create a TSV file for metadata from each organ
for (organ in organs) {
  
  filepath <- archr_filepaths[[organ]]
  atac_meta_path <- file.path(meta_dir, paste0("/", organ, "_atac_meta.tsv"))
  
  # Only process if the Seurat object exists, and the meta doesn't
  if (dir.exists(filepath) & !file.exists(atac_meta_path)) {
    
    message(paste0("Loading ArchRProject for ", organ, "..."))
    archr <- loadArchRProject(archr_filepaths[[organ]])
    archr_meta <- archr@cellColData %>% as.data.frame()
    write_tsv(archr_meta, atac_meta_path)
    
  } else message(organ, " meta already exists, skipping...")
  
}

```


<!-- Gather Per Sample Metadata -->

<!-- ```{r per_sample_meta} -->

<!-- # initialize dataframe -->
<!-- per_sample_meta <- data.frame(matrix(ncol = 7)) -->
<!-- colnames(per_sample_meta) <- c("Sample", "Organ", "Sex", "PCD", "PCW", "medianUMI", "medianGene") -->

<!-- for (organ in organs) { -->

<!--   rna_filepath <- file.path(meta_dir, organ) -->
<!--   rna_meta_path <- file.path(meta_dir, paste0("/", organ, "_rna_meta.tsv")) -->
<!--   atac_meta_path <- file.path(meta_dir, paste0("/", organ, "_atac_meta.tsv")) -->

<!--   # RNA metadata -->
<!--   if (file.exists(rna_meta_path)) { -->
<!--     message("reading rna metadata file for ", organ) -->

<!--     # use fread b/c it's faster -->
<!--     rna_meta <- data.table::fread(rna_meta_path, data.table = FALSE) -->

<!--     # fixes an issue when there is only one sex in a sample -->
<!--     if(rna_meta$sex %>% is.logical()) { -->
<!--       rna_meta$sex <- replace(rna_meta$sex, rna_meta$sex  == FALSE, "F") -->
<!--     } -->

<!--     # grab relevant metadata -->
<!--     rna_df <- rna_meta %>%  -->
<!--       dplyr::group_by(Sample) %>% -->
<!--       dplyr::summarise( -->
<!--         Organ = as.character(unique(organ)), -->
<!--         Sex = as.character(unique(sex)), -->
<!--         PCD = as.numeric(gsub("[^0-9]", "", unique(PCD))), -->
<!--         PCW = as.numeric(gsub("[^0-9]", "", unique(PCW))), -->
<!--         medianUMI = median(nCount_RNA), -->
<!--         medianGene = median(nFeature_RNA) -->
<!--       ) -->

<!--     # ATAC metadata -->
<!--     if (file.exists(atac_meta_path)) { -->

<!--       message("reading atac metadata file for ", organ) -->
<!--       atac_meta <- data.table::fread(atac_meta_path, data.table = FALSE) -->
<!--       atac_df <- atac_meta %>% -->
<!--         dplyr::group_by(Sample) %>% -->
<!--         dplyr::summarise( -->
<!--           medianFrag = median(nFrags), -->
<!--           medianTSS = median(TSSEnrichment) -->
<!--         ) -->

<!--       df <- left_join(rna_df, atac_df, by = "Sample") -->
<!--       per_sample_meta <- bind_rows(as_tibble(per_sample_meta), df) -->

<!--     }  -->
<!--     else { -->

<!--       message("no corresponding ATAC metadata available...") -->
<!--       rna.df$medianFrag <- NA -->
<!--       rna.df$medianTSS <- NA -->
<!--       per_sample_meta <- bind_rows(as_tibble(per_sample_meta), rna_df) -->
<!--     } -->
<!--   } -->
<!-- } -->

<!-- # remove the first row from initializing matrix -->
<!-- per_sample_meta <- per_sample_meta[-1,] -->

<!-- write_tsv(per_sample_meta, glue("{out}/per_sample_meta.tsv")) -->

<!-- ``` -->


# Per organ QC and metadata


```{r load_per_sample_meta}

per_sample_meta <- map_dfr(organ_codes, function(organ){

  filepath <- here::here(glue("output/01-preprocessing/02/shared/meta/{organ}_meta_sample.txt"))
  print(filepath)
  df <- read_tsv(filepath)
  print(dim(df))
  if(df$sex %>% is.logical()) {
    df$sex <- replace(df$sex, df$sex  == FALSE, "F") # Specify character type when only female sex in sample
  }
  df

})

DT::datatable(per_sample_meta)

write_tsv(per_sample_meta, glue("{out}/all_organs_keep_sample.tsv"))

print(head(per_sample_meta))

```



Plot frequency of each organ:

```{r pcw_dotplot, fig.width = 10, fig.hegiht = 5}

# timeline dotplot
per_sample_meta %>%
  mutate(organ = factor(organ, levels = rev(sort(names(cmap_organ))))) %>%
  ggplot(aes(x = pcw)) +
  geom_dotplot(aes(fill = organ, color = sex),
               binwidth = 0.1, binpositions = "all", stackgroups = TRUE, dotsize = 5, alpha = 0.8, width = 3, weight = 2) +
  scale_fill_manual(values = cmap_organ) +
  scale_color_manual(values = c("M" = "black", "F" = "white")) +
  scale_x_continuous(breaks = seq(10, 23)) +
  geom_hline(yintercept = 0) +
  theme(panel.grid = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        panel.border = element_blank())


```



Count samples per organ:

```{r barplot_freq_tissues, fig.width = 20, fig.height = 6}

# Plot barplot for each sample
p1 <- per_sample_meta %>%
  group_by(organ) %>%
  dplyr::count() %>%
  ggplot(aes(x = organ, y = n, fill = organ)) +
  geom_bar(stat = "identity") +
  theme_bw() +
  labs(x = "",
       y = "# Samples") +
  scale_fill_manual(values = cmap_organ) +
  scale_x_discrete(limits = rev) +
  coord_flip() +
  theme(
    panel.grid = element_blank(),
    axis.text = element_text(family = "sans", size = 10, color = "black"),
    axis.title = element_text(family = "sans", size = 11, color = "black"),
    legend.position = "none"
  )

# plot number of cells for each organ:
p2 <- per_sample_meta %>%
  dplyr::group_by(organ) %>%
  dplyr::summarise(
    total_ncells = sum(ncell)) %>%
  ggplot(aes(x = organ, y = total_ncells, fill = organ)) +
  geom_bar(stat = "identity") +
  theme_bw() +
  labs(x = "",
       y = "Count") +
  scale_fill_manual(values = cmap_organ) +
  scale_x_discrete(limits = rev) +
  scale_y_continuous(trans = scales::log10_trans(),
                     breaks = trans_breaks("log10", function(x) 10^x),
                     labels = trans_format("log10", math_format(10^.x))) +
  coord_flip() +
  theme(
    panel.grid = element_blank(),
    axis.text = element_text(family = "sans", size = 10, color = "black"),
    axis.title = element_text(family = "sans", size = 11, color = "black"),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "none"
  )

# Plot PCDs of each organ as a dotplot:
p3 <- ggplot(per_sample_meta, aes(x = pcd, y = organ, color = organ, shape = factor(sex))) +
  geom_point(size = 5, alpha = 0.8) +
  theme_bw() +
  labs(x = "Gestational age (days)",
       y = "") +
  scale_color_manual(values = cmap_organ) +
  scale_shape_manual(values = c(19, 17)) +
  scale_y_discrete(limits = rev) +
  theme(
    panel.grid = element_blank(),
    axis.text = element_text(family = "sans", size = 10, color = "black"),
    axis.title = element_text(family = "sans", size = 11, color = "black"),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom"
  )


# Plot RNA metrics as a box whisker plot:
p4 <- ggplot(per_sample_meta, aes(x = median_numi, y = organ, fill = organ)) +
  geom_boxplot(alpha = 0.8) +
  geom_jitter(height = 0.1) +
  scale_fill_manual(values = cmap_organ) +
  scale_x_log10() +
  scale_y_discrete(limits = rev) +
  theme_bw() +
  labs(x = "median # UMIs",
       y = "") +
  theme(
    panel.grid = element_blank(),
    axis.text = element_text(family = "sans", size = 10, color = "black"),
    axis.title = element_text(family = "sans", size = 11, color = "black"),
    legend.position = "none",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )

p5 <- ggplot(per_sample_meta, aes(x = median_ngene, y = organ, fill = organ)) +
  geom_boxplot(alpha = 0.8) +
  geom_jitter(height = 0.1) +
  scale_fill_manual(values = cmap_organ) +
  scale_x_log10() +
  scale_y_discrete(limits = rev) +
  theme_bw() +
  labs(x = "median # genes",
       y = "") +
  theme(
    panel.grid = element_blank(),
    axis.text = element_text(family = "sans", size = 10, color = "black"),
    axis.title = element_text(family = "sans", size = 11, color = "black"),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "none"
  )

# Plot ATAC metrics as a box whisker plot:
p6 <- ggplot(per_sample_meta, aes(x = median_nfrags, y = organ, fill = organ)) +
  geom_boxplot(alpha = 0.8) +
  geom_jitter(height = 0.1) +
  scale_fill_manual(values = cmap_organ) +
  scale_x_log10() +
  scale_y_discrete(limits = rev) +
  theme_bw() +
  labs(x = "median # fragments",
       y = NULL) +
  theme(
    panel.grid = element_blank(),
    axis.text = element_text(family = "sans", size = 10, color = "black"),
    axis.title = element_text(family = "sans", size = 11, color = "black"),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "none"
  )


p7 <- ggplot(per_sample_meta, aes(x = median_tss, y = organ, fill = organ)) +
  geom_boxplot(alpha = 0.8) +
  geom_jitter(height = 0.1) +
  scale_fill_manual(values = cmap_organ) +
  scale_y_discrete(limits = rev) +
  theme_bw() +
  labs(x = "median TSS enrichment",
       y = NULL) +
  theme(
    panel.grid = element_blank(),
    axis.text = element_text(family = "sans", size = 10, color = "black"),
    axis.title = element_text(family = "sans", size = 11, color = "black"),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "none"
  )


p8 <- ggplot(per_sample_meta, aes(x = median_frip, y = organ, fill = organ)) +
  geom_boxplot(alpha = 0.8) +
  geom_jitter(height = 0.1) +
  scale_fill_manual(values = cmap_organ) +
  scale_y_discrete(limits = rev) +
  theme_bw() +
  labs(x = "medianFRiP",
       y = NULL) +
  theme(
    panel.grid = element_blank(),
    axis.text = element_text(family = "sans", size = 10, color = "black"),
    axis.title = element_text(family = "sans", size = 11, color = "black"),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "none"
  )

plot_grid(p1, p2, p3, p4, p5, p6, p7, p8, nrow = 1, align = "h", axis = "tb")

# ggsave(paste0(figout, "/barplot_freq_tissues.pdf"), width=20, height=6)
```

Plot as 2D multiome metrics:

```{r dotPlot_multiome, fig.width = 15, fig.height = 5}

# nUMIs by nFrags
p1 <- ggplot(per_sample_meta, aes(x = median_numi, y = median_nfrags, color = organ)) +
  geom_point(size = 3, alpha = 0.9) +
  scale_x_continuous(trans = log10_trans(),
                     breaks = trans_breaks("log10", function(x) 10^x),
                     labels = trans_format("log10", math_format(10^.x))) +
  scale_y_continuous(trans = log10_trans(),
                     breaks = trans_breaks("log10", function(x) 10^x),
                     labels = trans_format("log10", math_format(10^.x))) +
  scale_color_manual(values = cmap_organ) +
  theme_bw() +
  theme(
    panel.grid = element_blank(),
    axis.text = element_text(family = "sans", size = 10, color = "black"),
    axis.title = element_text(family = "sans", size = 11, color = "black"),
    aspect.ratio = 1
    #legend.position = "none"
  )

# nGenes by FRiP
p2 <- ggplot(per_sample_meta, aes(x = median_ngene, y = median_frip, color = organ)) +
  geom_point(size = 3, alpha = 0.9) +
  scale_x_continuous(trans = log10_trans(),
                     breaks = trans_breaks("log10", function(x) 10^x),
                     labels = trans_format("log10", math_format(10^.x))) +
#  scale_y_continuous(trans = log10_trans(),
#                     breaks = trans_breaks("log10", function(x) 10^x),
#                     labels = trans_format("log10", math_format(10^.x))) +
  scale_color_manual(values = cmap_organ) +
  theme_bw() +
  theme(
    panel.grid = element_blank(),
    axis.text = element_text(family = "sans", size = 10, color = "black"),
    axis.title = element_text(family = "sans", size = 11, color = "black"),
    aspect.ratio = 1
    #legend.position = "none"
  )

# nGenes by TSS
p3 <- ggplot(per_sample_meta, aes(x = median_ngene, y = median_tss, color = organ)) +
  geom_point(size = 3, alpha = 0.9) +
  scale_x_continuous(trans = log10_trans(),
                     breaks = trans_breaks("log10", function(x) 10^x),
                     labels = trans_format("log10", math_format(10^.x))) +
  #  scale_y_continuous(trans = log10_trans(),
  #                     breaks = trans_breaks("log10", function(x) 10^x),
  #                     labels = trans_format("log10", math_format(10^.x))) +
  scale_color_manual(values = cmap_organ) +
  theme_bw() +
  theme(
    panel.grid = element_blank(),
    axis.text = element_text(family = "sans", size = 10, color = "black"),
    axis.title = element_text(family = "sans", size = 11, color = "black"),
    aspect.ratio = 1
    #legend.position = "none"
  )

plot_grid(p1, p2, p3, nrow = 1, align = "h", axis = "tb")

# ggsave(paste0(figout, "/dotPlot_multiome.pdf"), width=15, height=5)

```


# Per sample QC and metadata

Get cell metadata:

```{r cell_meta_rna, eval = FALSE}

# load all RNA cell meta into one table, add in the organ name
cell_meta_rna_paths <- list.files(here::here("output/02-global_analysis/01/per_cell_meta"), pattern = "*_rna_meta.tsv", full.names = TRUE)
cell_meta_rna <- map_dfr(cell_meta_rna_paths, ~ data.table::fread(.x, data.table = FALSE) %>%
                           add_column(Organ = str_split(basename(.x), "_", n = 3)[[1]][1], .before = 1))

# add organ code to cluster name
cell_meta_rna$Cluster <- map2_chr(cell_meta_rna$Organ, cell_meta_rna$Clusters,
                                  ~ paste0(organs_keep[organs_keep$organ == .x, ]$organcode, "_", .y))

# get a sample order by organ/age
sample_order <- cell_meta_rna %>%
  distinct(Sample, Organ, PCW) %>%
  arrange(Organ, PCW) %>%
  pull(Sample)

# n samples per organ
cell_meta_rna %>%
  distinct(Sample, Organ) %>%
  group_by(Organ) %>%
  dplyr::count()

# n cells per organ
cell_meta_rna %>%
  group_by(Organ) %>%
  dplyr::count()

dim(cell_meta_rna)

cell_meta_rna$Sample <- factor(cell_meta_rna$Sample, sample_order)

```



```{r cell_meta_atac, eval = FALSE}

# get cell barcodes in the df
# load all RNA cell meta into one table, add in the organ name
cell_meta_atac_paths <- list.files(here::here("output/02-global_analysis/01/per_cell_meta"), pattern = "*_atac_meta.tsv", full.names = TRUE)
cell_meta_atac <- map(cell_meta_atac_paths, ~ data.table::fread(.x, data.table = FALSE) %>%
                            tibble::rownames_to_column(var = "Barcode"))

cell_meta_atac <- map_dfr(cell_meta_atac, function(df) {

  if ("Clusters" %in% colnames(df)) df %>% dplyr::rename(RNA_Clusters = Clusters) %>% mutate(RNA_Clusters = as.character(RNA_Clusters))
  else df %>% mutate(RNA_Clusters = as.character(RNA_Clusters))

})

# get a cluster column matching RNA
per_sample_meta$Sample <- per_sample_meta$sample
cell_meta_atac <- cell_meta_atac %>%
  left_join(per_sample_meta, by = c("Sample")) %>%
  left_join(organs_keep, by = c("organ")) %>%
  mutate(Cluster = as.numeric(gsub("c", "", RNA_Clusters))) %>%
  mutate(Organ = organ) %>%
  arrange(Organ, Cluster) %>%
  mutate(Cluster = paste0(organ_code, "_", Cluster)) %>%
  mutate(Cluster = factor(Cluster, levels = unique(.$Cluster)))

```


Save metadata:

```{r save_meta, eval = FALSE}

save(cell_meta_atac, cell_meta_rna, file = glue("{out}/cell_meta_all.Rda"))

```

```{r}

load(glue("{out}/cell_meta_all.Rda"))

```


Basic stats:

```{r}

# cells
dim(cell_meta_rna)
dim(cell_meta_atac)

# samples
unique(cell_meta_rna$Sample)

```



<!-- # Testing -->

<!-- ```{r} -->

<!-- cell_meta_rna %>%  -->
<!--   filter(Organ == "Adrenal") %>%  -->
<!--   ggplot(aes(x = Sample)) + -->
<!--   geom_bar() -->

<!-- cell_meta_rna %>%  -->
<!--   filter(Organ == "Brain") %>%  -->
<!--   ggplot(aes(x = Sample)) + -->
<!--   geom_bar() -->

<!-- filter_and_plot <- function(my_organ) { -->

<!--   cell_meta_rna %>%  -->
<!--     filter(Organ == my_organ) %>%  -->
<!--     ggplot(aes(x = Sample)) + -->
<!--     geom_bar() -->

<!-- } -->


<!-- ``` -->


Let's generate a few plots for per-sample distributions of QC metrics:

## RNA

```{r n_cells_per_sample, fig.width = 13, fig.height = 5, cache = FALSE}

cell_meta_rna %>%
  group_by(Sample, Organ) %>%
  dplyr::count() %>%
  ggplot(aes(x = Sample, y = n)) +
  geom_col(aes(fill = Organ)) +
  geom_text(aes(label = n), angle = 90, hjust = -0.2) +
  ggtitle("Number of cells per sample") +
  scale_fill_manual(values = cmap_organ) +
  facet_grid(. ~ Organ, scales = "free_x", space = "free") +
  coord_cartesian(ylim = c(0, 30000)) +
  hide_legend() +
  rotate_x()

# ggsave(paste0(figout, "n_cells_per_sample.pdf"), width = 13, height = 5)
```

```{r n_UMIs_RNA_per_sample, fig.width = 13, fig.height = 5, cache = FALSE}

cell_meta_rna %>%
  ggplot(aes(x = Sample, y = nCount_RNA)) +
  geom_boxplot(aes(fill = Organ), outlier.shape = NA) +
  scale_fill_manual(values = cmap_organ) +
  rotate_x() +
  ggtitle("Number of UMIs per cell") +
  facet_grid(. ~ Organ, scales = "free_x", space = "free") +
  coord_cartesian(ylim = c(0, 20000)) +
  hide_legend()

# ggsave(paste0(figout, "n_UMIs_per_sample.pdf"), width = 13, height = 5)
```

```{r n_genes_RNA_per_sample, fig.width = 13, fig.height = 5, cache = FALSE}

cell_meta_rna %>%
  ggplot(aes(x = Sample, y = nFeature_RNA)) +
  geom_boxplot(aes(fill = Organ), outlier.shape = NA) +
  scale_fill_manual(values = cmap_organ) +
  rotate_x() +
  ggtitle("Number of genes per cell") +
  facet_grid(. ~ Organ, scales = "free_x", space = "free") +
  coord_cartesian(ylim = c(0, 7000)) +
  scale_y_continuous(breaks = seq(0, 10000, by = 1000)) +
  hide_legend()

# ggsave(paste0(figout, "n_genes_per_sample.pdf"), width = 13, height = 5)
```


## ATAC

```{r TSS_enrichment_ATAC_per_sample, fig.width = 13, fig.height = 5, cache = FALSE}

cell_meta_atac %>%
  ggplot(aes(x = Sample, y = TSSEnrichment)) +
  geom_boxplot(aes(fill = Organ), outlier.shape = NA) +
  geom_hline(yintercept = c(5, 10, 15), color = "red", linetype = 2) +
  scale_fill_manual(values = cmap_organ) +
  rotate_x() +
  ggtitle("TSS enrichment per cell") +
  facet_grid(. ~ Organ, scales = "free_x", space = "free") +
  coord_cartesian(ylim = c(0, 25)) +
  hide_legend()

# ggsave(paste0(figout, "TSS_enrichment_ATAC_per_sample.pdf"), width = 13, height = 5)

```

```{r n_fragments_ATAC_per_sample, fig.width = 13, fig.height = 5, cache = FALSE}

cell_meta_atac %>%
  ggplot(aes(x = Sample, y = nFrags)) +
  geom_boxplot(aes(fill = Organ), outlier.shape = NA) +
  geom_hline(yintercept = c(0, 2500, 5000, 7500), color = "red", linetype = 2) +
  scale_fill_manual(values = cmap_organ) +
  rotate_x() +
  ggtitle("Number of fragments per cell") +
  facet_grid(. ~ Organ, scales = "free_x", space = "free") +
  coord_cartesian(ylim = c(0, 30000)) +
  scale_y_continuous(breaks = seq(0, 30000, by = 2500)) +
  hide_legend()

# ggsave(paste0(figout, "n_fragments_ATAC_per_sample.pdf"), width = 13, height = 5)
```


# Per cluster QC and metadata

```{r breakdown_by_age_per_cluster, fig.width = 20, fig.height = 6, cache = FALSE}

cell_meta_rna %>%
  ggplot(aes(x = Cluster)) +
  geom_bar(aes(fill = PCW), position = "fill") +
  scale_fill_manual(values = cmap_pcw) +
  ggtitle("PCW") +
  rotate_x() +
  ggtitle("Cluster breakdown by age") +
  facet_grid(. ~ Organ, scales = "free_x", space = "free") +
  theme(legend.position = "bottom", panel.border = element_blank()) +
  guides(fill = guide_legend(nrow = 1))

# ggsave(paste0(figout, "breakdown_by_age_per_cluster.pdf"), width = 20, height = 6)
```


## RNA

```{r n_cells_per_cluster, fig.width = 20, fig.height = 5, cache = FALSE}

# get a cluster order
cell_meta_rna <- cell_meta_rna %>%
  arrange(Organ, Clusters) %>%
  mutate(Cluster = factor(Cluster, levels = unique(.$Cluster)))

cell_meta_rna %>%
  group_by(Cluster, Organ) %>%
  dplyr::count() %>%
  ggplot(aes(x = Cluster, y = n)) +
  geom_col(aes(fill = Organ)) +
  geom_text(aes(label = n), angle = 90, hjust = -0.2) +
  ggtitle("Number of cells per cluster") +
  scale_fill_manual(values = cmap_organ) +
  facet_grid(. ~ Organ, scales = "free_x", space = "free") +
  coord_cartesian(ylim = c(0, 30000)) +
  hide_legend() +
  rotate_x()

# ggsave(paste0(figout, "n_cells_per_cluster.pdf"), width = 20, height = 5)
```

```{r n_UMIs_RNA_per_cluster, fig.width = 20, fig.height = 5, cache = FALSE}

cell_meta_rna %>%
  ggplot(aes(x = Cluster, y = nCount_RNA)) +
  geom_boxplot(aes(fill = Organ), outlier.shape = NA) +
  geom_hline(yintercept = c(2500, 5000, 7500), color = "red", linetype = 2) +
  scale_fill_manual(values = cmap_organ) +
  rotate_x() +
  ggtitle("Number of UMIs per cell") +
  facet_grid(. ~ Organ, scales = "free_x", space = "free") +
  coord_cartesian(ylim = c(0, 25000)) +
  hide_legend()

# ggsave(paste0(figout, "n_UMIs_RNA_per_cluster.pdf"), width = 20, height = 5)
```

```{r n_genes_RNA_per_cluster, fig.width = 20, fig.height = 5, cache = FALSE}

cell_meta_rna %>%
  ggplot(aes(x = Cluster, y = nFeature_RNA)) +
  geom_boxplot(aes(fill = Organ), outlier.shape = NA) +
  geom_hline(yintercept = c(1000, 2000, 3000), color = "red", linetype = 2) +
  scale_fill_manual(values = cmap_organ) +
  rotate_x() +
  ggtitle("Number of genes per cell") +
  facet_grid(. ~ Organ, scales = "free_x", space = "free") +
  coord_cartesian(ylim = c(0, 7000)) +
  scale_y_continuous(breaks = seq(0, 10000, by = 1000)) +
  hide_legend()

# ggsave(paste0(figout, "n_genes_RNA_per_cluster.pdf"), width = 20, height = 5)
```


## ATAC

```{r TSS_enrichment_ATAC_per_cluster, fig.width = 20, fig.height = 5, cache = FALSE}

cell_meta_atac %>%
  ggplot(aes(x = Cluster, y = TSSEnrichment)) +
  geom_boxplot(aes(fill = Organ), outlier.shape = NA) +
  geom_hline(yintercept = c(5, 10, 15), color = "red", linetype = 2) +
  scale_fill_manual(values = cmap_organ) +
  rotate_x() +
  ggtitle("TSS enrichment per cell") +
  facet_grid(. ~ Organ, scales = "free_x", space = "free") +
  coord_cartesian(ylim = c(0, 25)) +
  hide_legend()

# ggsave(paste0(figout, "TSS_enrichment_ATAC_per_cluster.pdf"), width = 20, height = 5)
```

```{r n_fragments_ATAC_per_cluster, fig.width = 20, fig.height = 5, cache = FALSE}

cell_meta_atac %>%
  ggplot(aes(x = Cluster, y = nFrags)) +
  geom_boxplot(aes(fill = Organ), outlier.shape = NA) +
  geom_hline(yintercept = c(0, 2500, 5000, 7500), color = "red", linetype = 2) +
  scale_fill_manual(values = cmap_organ) +
  rotate_x() +
  ggtitle("Number of fragments per cell") +
  facet_grid(. ~ Organ, scales = "free_x", space = "free") +
  coord_cartesian(ylim = c(0, 30000)) +
  scale_y_continuous(breaks = seq(0, 30000, by = 2500)) +
  hide_legend()

# ggsave(paste0(figout, "n_fragments_ATAC_per_cluster.pdf"), width = 20, height = 5)
```


Total fragments per cluster:

```{r total_n_fragments_ATAC_per_cluster, fig.width = 20, fig.height = 5, cache = FALSE}

anno <- tibble::tribble(~Organ, ~y, ~Label,
                        "Adrenal", log10(2500000),  "2.5M",
                        "Adrenal", log10(5000000),  "5M",
                        "Adrenal", log10(10000000), "10M")

cell_meta_atac %>%
  dplyr::group_by(Organ, Cluster) %>%
  dplyr::summarize(Total_fragments = sum(nFrags)) %>%
  ggplot(aes(x = Cluster, y = log10(Total_fragments))) +
  geom_col(aes(fill = Organ)) +
  # label 2.5M, 5M, and 10M fragments
  geom_hline(yintercept = c(log10(10000000), log10(5000000), log10(2500000)), color = "red", linetype = 2) +
  geom_text(data = anno, aes(x = 1, y = y, label = Label), hjust = -1) +
  scale_fill_manual(values = cmap_organ) +
  rotate_x() +
  ggtitle("Log10 Total fragments per cluster") +
  facet_grid(. ~ Organ, scales = "free_x", space = "free") +
  hide_legend()

# ggsave(paste0(figout, "total_n_fragments_ATAC_per_cluster.pdf"), width = 20, height = 5)
```





# Compare to published fetal atlases

Collect metadata from Shendure Lab Fetal atlases:

```{r}

shendure_rna <- readr::read_csv(here::here("data/external/Domcke_Cao_metadata/ShendureLab_HumanFetal_RNA_Metadata.csv"))
shendure_atac <- readr::read_csv(here::here("data/external/Domcke_Cao_metadata/ShendureLab_HumanFetal_ATAC_Metadata.csv"))

# change case for ATAC metadata file to match
str_sub(shendure_atac$Tissue, 1, 1) <- str_sub(shendure_atac$Tissue, 1, 1) %>% str_to_upper()

shendure_rna$experiment <- "Cao et al."
shendure_atac$experiment <- "Domcke et al."
per_sample_meta$experiment <- "HDMA"

palette_experiment <- c("HDMA" = "forestgreen", "Cao et al." = "gray70", "Domcke et al." = "gray70")

# Filter organs
# Convert Cerebrum to Brain
shendure_rna$Organ[which(shendure_rna$Organ == "Cerebrum")] <- "Brain"
shendure_atac$Tissue[which(shendure_atac$Tissue == "Cerebrum")] <- "Brain"

# Convert Stomach to StomachEsophagus
shendure_rna$Organ[which(shendure_rna$Organ == "Stomach")] <- "StomachEsophagus"
shendure_atac$Tissue[which(shendure_atac$Tissue == "Stomach")] <- "StomachEsophagus"

filt.shendure_rna <- shendure_rna %>% filter(Organ %in% organs)
filt.shendure_atac <- shendure_atac %>% filter(Tissue %in% organs)

rna.meta.toPlot <- bind_rows(data.frame(Organ = per_sample_meta$organ,
                                        median_ngenes = per_sample_meta$median_ngene,
                                        median_numis = per_sample_meta$median_numi,
                                        experiment = per_sample_meta$experiment),
                             data.frame(Organ = filt.shendure_rna$Organ,
                                        median_ngenes = filt.shendure_rna$median_gene_count,
                                        median_numis = filt.shendure_rna$median_mRNA_count,
                                        experiment = filt.shendure_rna$experiment))

atac.meta.toPlot <- bind_rows(data.frame(Organ = per_sample_meta$organ,
                                         median_nfrags = per_sample_meta$median_nfrags,
                                         median_tss = per_sample_meta$median_tss,
                                         median_frip = per_sample_meta$median_frip,
                                         experiment = per_sample_meta$experiment),
                             data.frame(Organ = filt.shendure_atac$Tissue,
                                        median_nfrags = filt.shendure_atac$Median_total_fragments,
                                        median_tss = filt.shendure_atac$TSS_enrichment,
                                        median_frip = filt.shendure_atac$Median_per_cell_FRiP,
                                        experiment = filt.shendure_atac$experiment))

```

Plot QC comparisons:

```{r BoxPlot_HDMA_vs_Cao_RNA, fig.width = 6, fig.heigth = 5}

p1 <- ggplot(rna.meta.toPlot, aes(x = Organ, y = median_numis, fill = experiment)) +
  geom_boxplot(outlier.shape = NA) +
  scale_fill_manual(values = palette_experiment, name = "Study") +
  scale_y_log10() +
  theme_bw() +
  theme(axis.text.x = element_blank()) +
  xlab("") +
  ylab("median # UMIs")

p2 <- ggplot(rna.meta.toPlot, aes(x = Organ, y = median_ngenes, fill = experiment)) +
  geom_boxplot(outlier.shape = NA) +
  scale_fill_manual(values = palette_experiment, name = "Study") +
  scale_y_log10() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  xlab("") +
  ylab("median # genes")

plot_grid(p1, p2, ncol = 1, rel_heights = c(0.45, 0.55), align = "v", axis = "rl")

# ggsave(paste0(figout, "BoxPlot_HDMA_vs_Cao_RNA.pdf"), width = 6, height = 5)
```

```{r boxPlot_HDMA_vs_Cao_ATAC, fig.width = 6, fig.height = 5}

p1 <- ggplot(atac.meta.toPlot, aes(x = Organ, y = median_nfrags, fill = experiment)) +
  geom_boxplot(outlier.shape = NA) +
  scale_fill_manual(values = palette_experiment, name = "Study") +
  scale_y_log10() +
  theme_bw() +
  theme(axis.text.x = element_blank()) +
  xlab("") +
  ylab("median # fragments")

# p2 <- ggplot(atac.meta.toPlot, aes(x = Organ, y = median_frip, fill = experiment)) +
#   geom_boxplot(outlier.shape = NA) +
#   scale_fill_manual(values = palette_experiment, name = "Study") +
#   scale_y_log10() +
#   theme_bw() +
#   theme(axis.text.x = element_blank()) +
#   xlab("") +
#   ylab("median FRiP")

p3 <- ggplot(atac.meta.toPlot, aes(x = Organ, y = median_tss, fill = experiment)) +
  geom_boxplot(outlier.shape = NA) +
  scale_fill_manual(values = palette_experiment, name = "Study") +
  scale_y_log10() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  xlab("") +
  ylab("median TSS enrichment")

plot_grid(p1, p3, ncol = 1, rel_heights = c(0.44, 0.55), align = "v", axis = "rl")

# ggsave(paste0(figout, "boxPlot_HDMA_vs_Cao_ATAC.pdf"), width = 6, height = 5)
```



# Session info

```{r sinfo, cache = FALSE}

.libPaths()
sessionInfo()

```

