---
title: "Filter global peaks"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: 
    code_folding: show
    theme: flatly
    highlight: pygments
    toc: yes
    toc_float:
      collapsed: true
    toc_depth: 4
    number_sections: yes
    keep_md: no
---

# Filter global peak set

## Read all global peaks

```{r}
library(tidyverse)
library(here)
library(GenomicRanges)

out <- here("output/04-enhancers/01")

all_peaks <- readRDS(paste0(here("output/01-preprocessing/03"), "/peaks_gr.rds") )
length(all_peaks)
all_peaks %>% head
```

Add peak name in the format chromosome_start_end

```{r}
all_peaks$peak_name <- paste0(seqnames(all_peaks), "_", ranges(all_peaks) %>% start, "_",ranges(all_peaks) %>% end)

all_peaks %>% head
```

Convert GRanges into a 1-based data frame

```{r}
peaks_df <- data.frame(all_peaks)
peaks_df %>% head
```

Filter out columns we dont need

```{r}
rownames(peaks_df) <- peaks_df$peak_name 
peaks_df <- peaks_df[!colnames(peaks_df) %in% c("width", "strand","idx", "N")]
peaks_df %>% head
```

Test if we can easily convert the dataframe back to a GRanges object

```{r}
GRanges(peaks_df) %>% head
```

Save data frame with and without the ArchR default annotation columns

```{r}
saveRDS(peaks_df, paste0(out, "/peaks_all_df_annots.rds")) # 1-based
saveRDS(peaks_df[c("seqnames", "start", "end")], paste0(out, "/peaks_all_df.rds")) # 1-based

# also export as 0-based bed file
rownames(peaks_df) <- NULL
rtracklayer::export.bed(peaks_df, here::here("output/04-enhancers/01/hdma_global_acCREs.bed")) # 0-based
```

## Filter for non-exonic peaks only

```{r}
all_peaks$peakType %>% table

peaks_noexon <- all_peaks[all_peaks$peakType != "Exonic"]
peaks_noexon$peakType %>% table

length(peaks_noexon)
peaks_noexon %>% head
```

Add peak name in the format chromosome_start_end

```{r}
peaks_noexon$peak_name <- paste0(seqnames(peaks_noexon), "_", ranges(peaks_noexon) %>% start, "_",ranges(peaks_noexon) %>% end)

peaks_noexon %>% head
```

Explore some other columns in the table

```{r}
peaks_noexon$idx %>% quantile
```

```{r}
peaks_noexon$N %>% quantile
```

```{r}
peaks_noexon$score %>% quantile
```

```{r}
peaks_noexon$replicateScoreQuantile %>% quantile
```

Convert GRanges into a data frame

```{r}
peaks_df <- data.frame(peaks_noexon)
peaks_df %>% head
```

Filter out columns we dont need

```{r}
rownames(peaks_df) <- peaks_df$peak_name 
peaks_df <- peaks_df[!colnames(peaks_df) %in% c("width", "strand","idx", "N")]
peaks_df %>% head
```

Test if we can easily convert the dataframe back to a GRanges object

```{r}
GRanges(peaks_df) %>% head
```

Save data frame with and without the ArchR default annotation columns

```{r}
saveRDS(peaks_df, paste0(out, "/peaks_noexon_df_annots.rds")) # 1-based
saveRDS(peaks_df[c("seqnames", "start", "end")], paste0(out, "/peaks_noexon_df.rds")) # 1-based
```
