---
title: "ATAC co-accessibility analysis"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: 
    code_folding: show
    theme: flatly
    highlight: pygments
    toc: yes
    toc_float:
      collapsed: true
    toc_depth: 4
    number_sections: yes
    keep_md: no
---

```{r config, warning = FALSE}

# **MODIFY THIS CHUNK**
project_id  <- "HDMA-public" # determines the name of the cache folder
doc_id      <- "04-enhancers/04" # determines name of the subfolder of `figures` where pngs/pdfs are saved
out         <- here::here("output/", doc_id); dir.create(out, recursive = TRUE)
figout      <- here::here("figures/", doc_id, "/")
cache       <- paste0("~/scratch/cache/", project_id, "/", doc_id, "/")

```

```{r setup, include = FALSE}

# NO NEED TO MODIFY THIS CHUNK

knitr::knit_hooks$set( echo_label = function(before, options, envir) {
  if ( before ) {
    # Do nothing
  } else sprintf('<br><span style="color:#1abc9c">~[output @ *%s*]~</span>',
                 paste0(options$fig.path, "/", options$label, "...") )
})

knitr::opts_chunk$set(message = TRUE,
                      warning = FALSE,
                      error = FALSE,
                      cache = FALSE,
                      cache.path = cache,
                      fig.path = figout,
                      fig.keep = "all",
                      dev = c("png", "pdf"),
                      cache.lazy = FALSE)

grDevices::pdf.options(useDingbats = FALSE)

options(knitr.table.format = "html") 
knitr::opts_knit$set(width = 1600)

set.seed(100)

```

# calculate the number of coaccessible peaks from global peak set

```{r}
suppressPackageStartupMessages({
  library(ArchR)
  library(tidyverse)
  library(here)
})

# Get additional functions, etc.:
scriptPath <- here("code/utils")
source(paste0(scriptPath, "/plotting_config.R"))
source(paste0(scriptPath, "/misc_helpers.R"))
source(paste0(scriptPath, "/matrix_helpers.R"))
source(paste0(scriptPath, "/archr_helpers.R"))
source(paste0(scriptPath, "/hdma_palettes.R"))


```

## read project
```{r}
global_obj <- here("output/01-preprocessing/03/allSamples/")
proj <- loadArchRProject(global_obj)
proj
```

```{r}
coacc <- getCoAccessibility(proj, returnLoops = F)
peaks <- metadata(coacc)$peakSet
head(coacc)
head(peaks)
```

## convert peak ID to peak ranges
```{r}
coacc_tab <- table(coacc$queryHits) %>% as.data.frame %>% 
  dplyr::rename(peakID=Var1, n_coaccessible_peaks=Freq)
head(coacc_tab)
peak_hits <- peaks[coacc_tab$peakID %>% as.character %>% as.integer,]
head(peak_hits)
```

## adding peak names from ranges
```{r}
coacc_tab$peak_name <- paste0(seqnames(peak_hits), "_", ranges(peak_hits) %>% start, "_",ranges(peak_hits) %>% end)
rownames(coacc_tab) <- coacc_tab$peak_name
#coacc_tab$peakID  <- NULL 
head(coacc_tab)
```
## load all global peaks and double check that all peaknames here are present in global peakset
```{r}
global <- readRDS(here::here("output/04-enhancers/01/peaks_all_df_annots.rds"))
sum(rownames(coacc_tab) %in% rownames(global)) 
sum(!rownames(coacc_tab) %in% rownames(global)) 
```

## save output
```{r}
saveRDS(coacc_tab, paste0(out, "/n_coacc_peaks_df.rds"))
```

## read output
```{r}
coacc_tab <- readRDS(paste0(out, "/n_coacc_peaks_df.rds"))
```

## some summary stats and plots
### distribution of n_coaccessible_peaks
```{r jplot_ncoacc_peak, fig.width=5, fig.height=5}
quantile(coacc_tab$n_coaccessible_peaks)
quantile(coacc_tab$n_coaccessible_peaks, seq(0, 1, by = 0.1))
coacc_tab_sorted <- coacc_tab %>% dplyr::arrange(n_coaccessible_peaks)
coacc_tab_sorted$rank <- 1:dim(coacc_tab_sorted)[1]
head(coacc_tab_sorted)

p <- ggplot(coacc_tab_sorted, aes(x=rank, y=n_coaccessible_peaks)) + geom_point() + theme_BOR() 
h <- 10
p1 <- p + geom_hline(yintercept=h, col="red", linetype="dashed") +
  geom_text(aes(0,h),label = paste0("y=",h), vjust = -1, col="red")

h <- 20
p2 <- p + geom_hline(yintercept=h, col="red", linetype="dashed") +
  geom_text(aes(0,h),label = paste0("y=",h), vjust = -1, col="red")

h <- 15
p3 <- p + geom_hline(yintercept=h, col="red", linetype="dashed") +
  geom_text(aes(0,h),label = paste0("y=",h), vjust = -1, col="red")


# pdf(paste0(figout, "/jplot_ncoacc_peak.pdf"), width=5, height=5)
p1  
p2
p3
# dev.off()
```

### distribution of n_coaccessible_peaks > 1
```{r}
npeaks <- coacc_tab$n_coaccessible_peaks
length(npeaks)
length(npeaks[npeaks > 1])
quantile(npeaks[npeaks > 1])
```
### inspect peak types in top coaccessible peaks
```{r peaktype_topaccpeaks}
coacc_cutoff <- 15
#coacc_top <- coacc_tab %>% dplyr::filter(n_coaccessible_peaks > quantile(coacc_tab$n_coaccessible_peaks, 0.9))
coacc_top <- coacc_tab %>% dplyr::filter(n_coaccessible_peaks > coacc_cutoff)
coacc_top <- merge(coacc_top, global, by.x=0, by.y=0) %>% column_to_rownames(var="Row.names")

ggplot(coacc_top,aes(x=peakType, fill=peakType)) + geom_bar()  + 
    scale_fill_manual(values = c("#75B76C", "#8DBFE6", "#5B328C", "#F5C768")) + theme_BOR()

```
### also inspect peak types in the targets of top coaccessible peaks
```{r peaktype_topaccpeaks_targets}
peak_hits <- peaks[coacc$queryHits,]
coacc$peak_name <- paste0(seqnames(peak_hits), "_", ranges(peak_hits) %>% start, "_",ranges(peak_hits) %>% end)

# filter coacc links for top coaccessible peaks
coacc_toplinks <- coacc[coacc$peak_name %in% rownames(coacc_top),]
target_peaks <- peaks[coacc_toplinks$subjectHits,]
coacc_toplinks$subject_peak_name <- paste0(seqnames(target_peaks), "_", ranges(target_peaks) %>% start, "_",ranges(target_peaks) %>% end)
rownames(coacc_toplinks) <- coacc_toplinks$subject_peak_name
coacc_toplinks <- coacc_toplinks %>% as.data.frame
coacc_toplinks$subject_peak_type <- global[coacc_toplinks$subject_peak_name, "peakType"]

ggplot(coacc_toplinks,aes(x=subject_peak_type, fill=subject_peak_type)) + geom_bar()  + 
    scale_fill_manual(values = c("#75B76C", "#8DBFE6", "#5B328C", "#F5C768")) + theme_BOR()

ggplot(global,aes(x=peakType, fill=peakType)) + geom_bar()  + 
    scale_fill_manual(values = c("#75B76C", "#8DBFE6", "#5B328C", "#F5C768")) + theme_BOR()

```

### organ origin of the top peaks with most number of coaccessible peaks
```{r boxplot_topaccpeak_origin, fig.width=10, fig.height=6}
organ_stats <- sapply(coacc_top$GroupReplicate, function(n){strsplit(n, split="_")[[1]][1]}) %>% 
  table %>% as.data.frame %>% dplyr::rename(code=".", n_peaks_top10_coaccessible="Freq")

organ_allpeak_stats <- sapply(global$GroupReplicate, function(n){strsplit(n, split="_")[[1]][1]}) %>% 
  table %>% as.data.frame %>% dplyr::rename(code=".", n_peaks_total="Freq")

organ_stats <- merge(organ_stats, organ_allpeak_stats)
organ_stats$ratio_top10_accessible <- organ_stats$n_peaks_top10_coaccessible / organ_stats$n_peaks_total
head(organ_stats)

# convert organ code to organ name for color palettes
toKeep <- read_tsv(here::here("code/02-global_analysis/01-organs_keep.tsv")) %>% as.data.frame
rownames(toKeep) <- toKeep$organcode
organ_stats$organ_name <- toKeep[organ_stats$code, "organ"]
organ_stats$color <- cmap_organ[organ_stats$organ_name]
head(organ_stats)

# pdf(paste0(figout,"/boxplot_n_coacc_peak_organ.pdf"), width=10, height=6)
ggplot(organ_stats, aes(x=organ_name, y=n_peaks_total, fill=organ_name)) + geom_col() + theme_BOR() + scale_fill_manual(values = organ_stats$color) + theme(axis.text.x = element_text(angle=90, hjust = 1)) + 
  ylab("total # of global peaks \noriginating from organ")
ggplot(organ_stats, aes(x=organ_name, y=n_peaks_top10_coaccessible, fill=organ_name)) + geom_col() + theme_BOR() + scale_fill_manual(values = organ_stats$color) + theme(axis.text.x = element_text(angle=90, hjust = 1)) + ylab("# of organ-originated peaks \nin highly coaccessible peaks")

ggplot(organ_stats, aes(x=organ_name, y=ratio_top10_accessible*100, fill=organ_name)) + geom_col() + theme_BOR() + scale_fill_manual(values = organ_stats$color) + theme(axis.text.x = element_text(angle=90, hjust = 1)) + ylab("% of organ-originated peaks \nin highly coaccessible peaks")

# dev.off()
```

### plot tracks for top genes near the peaks with the highest number of coaccesible peaks 
```{r tracks_highly_coacc_genes_n15, fig.width=5, fig.height=5, eval=F}
# hcag = highly co-accessible gene
hcag <- coacc_top %>% arrange(desc(n_coaccessible_peaks)) %>% dplyr::select(nearestGene) %>% table %>% sort(decreasing=T) %>% head(30)

proj$organ_code <- sapply(proj$RNA_Clusters, function(n){strsplit(n, split="_")[[1]][1]})

p <- plotBrowserTrack(
  ArchRProj = proj, 
  groupBy = "organ_code", 
  geneSymbol = hcag %>% names, 
  upstream = 100000,
  downstream = 100000,
  loops = getCoAccessibility(proj)
)

saveRDS(hcag, paste0(out,"/highly-coaccessible-genes.rds"))

# plotPDF(plotList = p,
#         name = "plot_tracks_highly_coaccessible_genes_n15.pdf",
#         ArchRProj = proj,
#         addDOC = FALSE, width = 5, height = 5)
p

```

### get all cluster annotations

```{r get_annot}
# add organ colors
col.data.frame <- data.frame(Color=cmap_organ) %>% dplyr::mutate(organ=names(cmap_organ))

organ.code.list <- toKeep$organcode
all.annots <- lapply(1:length(organ.code.list), function(i){
  # read cluster id annotation
  annot <- read.csv(sprintf(here("output/01-preprocessing/02/shared/meta/%s_meta_cluster.txt"), organ.code.list[i]), sep="\t") %>% as.data.frame
  rownames(annot) <- annot$cb
  annot$L0_clusterID <- paste0(organ.code.list[i],"_",annot$L1_clusterID)
  annot$L3_clusterID <- paste0(organ.code.list[i],"_",annot$L3_clusterName)
  return(annot)
})
all.annots <- dplyr::bind_rows(all.annots)
rownames(all.annots) <- all.annots$L0_clusterID

all.annots <- all.annots %>%
  mutate(Cluster = paste0(organ_code, "_",L1_clusterID)) %>%
  mutate(Cluster_labelled = paste0(organ_code, "_", L1_clusterID, "_", L3_clusterName)) %>%
  left_join(col.data.frame) %>%
  tibble::column_to_rownames(var = "Cluster")
dim(all.annots)
head(all.annots)
```


### plot tracks for top genes by cell type, subset to the organ of interest
```{r tracks_top_coaccgene, fig.width=10, fig.height=6, eval=F}
proj$organ_code <- sapply(proj$RNA_Clusters, function(n){strsplit(n, split="_")[[1]][1]})
proj$cell_type <- all.annots[proj$RNA_Clusters, "L3_clusterName"]

plot_track_goi <- function(organ_code, gene, proj){
  p <- plotBrowserTrack(
  ArchRProj = proj[proj$organ_code==organ_code], 
  groupBy = "cell_type", 
  geneSymbol = gene, 
  upstream = 100000,
  downstream = 100000,
  loops = getCoAccessibility(proj)
  )

  # pdf(file.path(out, sprintf("%s/plot-browser-track-%s-%s.pdf", figout, organ_code, gene)), height=6, width=6)
  plot(p[[gene]])
  # dev.off()
}


plot_track_goi("EY", "OTX2", proj)

plot_track_goi("HT", "NKX2-5", proj)
plot_track_goi("SP", "NKX2-5", proj)

plot_track_goi("HT", "GATA4", proj)

plot_track_goi("HT", "GATA6", proj)
plot_track_goi("AG", "GATA6", proj)

plot_track_goi("LI", "APOB", proj)

plot_track_goi("MU", "MYH3", proj)
plot_track_goi("TR", "MYH3", proj)

plot_track_goi("TM", "IKZF1", proj)
plot_track_goi("LI", "IKZF1", proj)

plot_track_goi("TM", "CD8A", proj)


```

