---
title: "Explore VISTA enhancer database"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: 
    code_folding: show
    theme: flatly
    highlight: pygments
    toc: yes
    toc_float:
      collapsed: true
    toc_depth: 4
    number_sections: yes
    keep_md: no
---

```{r config, warning = FALSE}

# **MODIFY THIS CHUNK**
project_id  <- "HDMA-public" # determines the name of the cache folder
doc_id      <- "04-enhancers/08" # determines name of the subfolder of `figures` where pngs/pdfs are saved
out         <- here::here("output/", doc_id); dir.create(out, recursive = TRUE)
figout      <- here::here("figures/", doc_id, "/")
cache       <- paste0("~/scratch/cache/", project_id, "/", doc_id, "/")
script_path <- here::here("code/utils/")

```

```{r setup, include = FALSE}

# NO NEED TO MODIFY THIS CHUNK

knitr::knit_hooks$set( echo_label = function(before, options, envir) {
  if ( before ) {
    # Do nothing
  } else sprintf('<br><span style="color:#1abc9c">~[output @ *%s*]~</span>',
                 paste0(options$fig.path, "/", options$label, "...") )
})

knitr::opts_chunk$set(message = TRUE,
                      warning = FALSE,
                      error = FALSE,
                      cache = FALSE,
                      cache.path = cache,
                      fig.path = figout,
                      fig.keep = "all",
                      dev = c("png", "pdf"),
                      cache.lazy = FALSE)

grDevices::pdf.options(useDingbats = FALSE)

options(knitr.table.format = "html") 
knitr::opts_knit$set(width = 1600)

set.seed(100)

```

***

# Overview

We received the VISTA-validated enhancer set from collaborators. Here we explore the
data to guide our handling of the dataset for training/testing validation.


# Set up

```{r libs, warning = FALSE, message = FALSE, echo = TRUE, cache = FALSE}

# libraries
library(here)
library(dplyr)
library(tidyr)
library(ggplot2)
library(purrr)
library(glue)
library(readr)
library(tibble)
library(cowplot)
library(GenomicRanges)
library(rtracklayer)

# shared project scripts/functions
source(file.path(script_path, "plotting_config.R"))
source(file.path(script_path, "hdma_palettes.R"))
source(file.path(script_path, "sj_scRNAseq_helpers.R"))

ggplot2::theme_set(theme_BOR())

```


# Load data

```{r load_dict}

tissue_dict <- read_csv(here::here("data/external/2024-01-24_vista_@mkosicki/tissue.dictionary.csv")) %>% 
  mutate(HDMA_tissue = case_when(
    tissue.short == "eye" ~ "Eye",
    tissue.short == "forebrain" ~ "Brain",
    tissue.short == "heart" ~ "Heart",
    tissue.short == "melanocytes" ~ "Skin",
    tissue.short == "liver" ~ "Liver",
    TRUE ~ "None"
  ))
DT::datatable(tissue_dict)

# make a palette
cmap_vista_tissue                <- rep("gray70", nrow(tissue_dict))
names(cmap_vista_tissue)         <- tissue_dict$tissue.short
cmap_vista_tissue["heart"]       <- cmap_organ["Heart"]
cmap_vista_tissue["forebrain"]   <- cmap_organ["Brain"]
cmap_vista_tissue["liver"]       <- cmap_organ["Liver"]
cmap_vista_tissue["melanocytes"] <- cmap_organ["Skin"]
cmap_vista_tissue["eye"]         <- cmap_organ["Eye"]

```


```{r load_vista}

vista <- read_tsv(here::here("data/external/2024-01-24_vista_@mkosicki/2024-01-24_vista.clean.tsv"))
tibble::glimpse(vista)

unique(vista$assembly)
sum(is.na(vista$vista_id))

length(unique(vista$vista_id))
length(unique(vista$coord))

vista %>% 
  group_by(coord, assembly) %>% 
  summarize(assemblies = length(unique(assembly)))

```



Tissue positivity is represented as follows:

> "tissue", "tissue_positive" and "denominator columns" provide an estimate of how reliable given tissue annotation is
- e.g. tissue=fb;mb, tissue_positive=5;9, denominator=9 should be read as
"5/9 transgenic embryos were positive for forebrain, 9/9 for midbrain", which implies
midbrain staining was stronger and/or more reproducible than forebrain staining

```{r process_vista}

# define a function to split out the tissue positivity into a tidy dataframe
process_validation <- function(vista_id, tissue, tissue_positive, denominator) {
  
  if (tissue == "neg") {
    
    tissues <- "none"
    positives <- 0
    positives_prop <- 0
    
  } else {
    
    tissues <- stringr::str_split(tissue, ";")[[1]]
    positives <- as.numeric(stringr::str_split(tissue_positive, ";")[[1]])
    positives_prop <- positives / denominator
    
  }
  
  return(data.frame("vista_id"      = vista_id,
                    "tissue"        = tissues,
                    "n_positive"    = positives,
                    "denominator"   = denominator,
                    "prop_positive" = positives_prop))
  
}

vista_long <- map_dfr(1:nrow(vista),
                      ~ process_validation(vista$vista_id[.x],
                                           vista$tissue[.x],
                                           vista$tissue_positive[.x],
                                           vista$denominator[.x])) %>% 
  dplyr::rename(tissue.abb = tissue) %>% 
  left_join(tissue_dict, by = "tissue.abb")

write_tsv(vista_long, glue("{out}/vista_long_with_atlas_tissues.tsv"))

head(vista_long)

```



# Exploratory plots


```{r tissue_positive_stats, fig.width = 9, fig.height = 6}

vista_long %>% 
  filter(tissue != "none") %>% 
  ggplot(aes(x = prop_positive)) +
  geom_histogram(bins = 10, aes(fill = tissue.short)) +
  scale_fill_manual(values = cmap_vista_tissue) +
  facet_wrap(~ tissue.short) +
  hide_legend() +
  theme(panel.grid.major.y = element_line(),
        panel.grid.minor.y = element_line())

vista_long %>% 
  filter(tissue != "none") %>% 
  ggplot(aes(x = prop_positive)) +
  geom_histogram(bins = 10, aes(fill = tissue.short)) +
  scale_fill_manual(values = cmap_vista_tissue) +
  facet_wrap(~ tissue, scales = "free_y") +
  hide_legend() +
  theme(panel.grid.major.y = element_line())

vista_long %>% 
  filter(tissue != "none") %>% 
  ggplot(aes(x = prop_positive)) +
  stat_ecdf(geom = "step", aes(color = tissue.short)) +
  scale_color_manual(values = cmap_vista_tissue) +
  facet_wrap(~ tissue.short, scales = "free_y") +
  hide_legend()

```


```{r tissue_positive_0.3, fig.width = 8, fig.height = 8}

vista_long %>% 
  filter(tissue != "none") %>% 
  filter(prop_positive > 0.3) %>% 
  ggplot(aes(x = tissue)) +
  geom_bar(aes(fill = tissue.short)) +
  scale_fill_manual(values = cmap_vista_tissue) +
  hide_legend() +
  theme(panel.grid.major.y = element_line(),
        panel.grid.minor.y = element_line()) +
  ggtitle("# of elements positive in >30% of samples") +
  coord_flip()


```

Figure out what's contained in the "other" tissue annotations, by looking at 
the external notes:

<details>

```{r check_other}

vista %>%
  filter(grepl("other", tissue)) %>% 
  pull(external_note) %>% 
  table() %>% 
  sort(decreasing = TRUE)

```

</details>


# Session info

```{r sinfo, cache = FALSE}

.libPaths()
sessionInfo()

```

