---
title: "Exploration of de novo learned motifs"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: 
    code_folding: show
    theme: flatly
    highlight: pygments
    toc: yes
    toc_float:
      collapsed: true
    toc_depth: 4
    number_sections: yes
    keep_md: no
---



```{r config, warning = FALSE, message = FALSE}

# **MODIFY THIS CHUNK**
library(here)
kundaje_dir <- trimws(readr::read_lines(here("code/AK_PROJ_DIR.txt")))
doc_id      <- "01"
out         <- here("output/03-chrombpnet/03-syntax/", doc_id); dir.create(out, recursive = TRUE)
figout      <- here("figures/03-chrombpnet/03-syntax", doc_id, "/"); dir.create(figout, recursive = TRUE)
chrombpnet_dir <- here("output/03-chrombpnet")

```


```{r setup, include = FALSE}

# NO NEED TO MODIFY THIS CHUNK

knitr::knit_hooks$set( echo_label = function(before, options, envir) {
  if ( before ) {
    # Do nothing
  } else sprintf('<br><span style="color:#1abc9c">~[output @ *%s*]~</span>',
                 paste0(options$fig.path, "/", options$label, "...") )
})

knitr::opts_chunk$set(message = TRUE,
                      warning = FALSE,
                      error = FALSE,
                      cache = FALSE,
                      fig.path = figout,
                      fig.keep = "all",
                      dev = c("png", "pdf"),
                      cache.lazy = FALSE)

grDevices::pdf.options(useDingbats = FALSE)

options(knitr.table.format = "html") 
knitr::opts_knit$set(width = 1600)

set.seed(100)

```

***

# Overview

In this notebook, we do global investigations of the de novo motifs:

- computing stats on the motif instances such as frequencies per peak
- summarize each motif's distribution of distances to nucleosome dyads and peak flanks
- distribution of motif instances in genomic features
- distribution of motif instances in tissues and tissue compartments

# Set up

```{r libs, warning = FALSE, message = FALSE, echo = TRUE, cache = FALSE}

library(glue)
library(scales)
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(purrr)
library(stringr)
library(ggrepel)
library(cowplot)
library(plotly)
library(ggseqlogo)
library(BPCells)

script_path <- here("code/utils/")
source(file.path(script_path, "plotting_config.R"))
source(file.path(script_path, "hdma_palettes.R"))
source(file.path(script_path, "sj_scRNAseq_helpers.R"))
source(file.path(script_path, "chrombpnet_utils.R"))

ggplot2::theme_set(theme_BOR())

```


# Load & wrangle data

## Load cluster, motif, and hits data

Cluster metadata:

```{r load_cluster_meta}

cluster_meta <- read_csv(here("output/05-misc/03/TableS2_cluster_meta_qc.csv"))

chrombpnet_models_keep <- read_tsv(here("output/03-chrombpnet/01-models/qc/chrombpnet_models_keep2.tsv"),
                                   col_names = c("Cluster", "Folds_keep", "Cluster_ID"))

cluster_order <- cluster_meta %>% arrange(organ, L1_annot) %>% pull(Cluster_ChromBPNet)
organs <- unique(cluster_meta$organ) %>% sort()
organs[organs == "StomachEsophagus"] <- "Stomach"

head(chrombpnet_models_keep)

```

Load metrics:

```{r}

chrombpnet_metrics <- read_tsv(here("output/03-chrombpnet/01-models/qc/chrombpnet_metrics.tsv")) %>% 
  dplyr::select(Cluster_ChromBPNet, total_frags) %>% 
  distinct()

```


Load number of peaks:

```{r load_npeaks}

npeaks <- read_table(here("output/03-chrombpnet/01-models/qc/npeaks.tsv"),
                     col_names = c("n_peaks", "peak_file")) %>% 
  filter(peak_file != "total")

nrow(npeaks) == 203

npeaks <- npeaks %>% 
  rowwise() %>% 
  mutate(peak_file = basename(peak_file)) %>% 
  separate(peak_file, into = c("Cluster_ChromBPNet", "drop"), sep = "__") %>% 
  dplyr::select(-drop)

cluster_meta <- cluster_meta %>%
  left_join(npeaks, by = "Cluster_ChromBPNet") %>% 
  left_join(chrombpnet_metrics, by = "Cluster_ChromBPNet")

```


Load merged modisco reports:

```{r load_modisco_merged}

modisco_merged <- read_tsv(here("output/03-chrombpnet/02-compendium/modisco_compiled_anno/modisco_merged_reports.tsv")) %>% 
  # get short cluster ID
  left_join(cluster_meta %>% dplyr::select(Cluster, Cluster_ChromBPNet),
            by = c("component_celltype" = "Cluster_ChromBPNet"))

head(modisco_merged)

```


Load annotated patterns:

```{r load_motifs_compiled}

# manually annotated motif patterns
motifs_compiled_all <- read_tsv(here("code/03-chrombpnet/02-compendium/04d-ChromBPNet_de_novo_motifs.tsv")) %>% 
  mutate(motif_name = paste0(idx_uniq, "|", annotation)) %>% 
  dplyr::relocate(motif_name, .after = pattern) %>%
  dplyr::relocate(idx_uniq, .after = idx) %>% 
  # rename categories
  mutate(category = ifelse(annotation == "exclude", "exclude", category),
         category = case_match(category,
                               "composite_homo" ~ "homocomposite",
                               "composite_hetero" ~ "heterocomposite",
                               .default = category),
         category = factor(category, levels = names(cmap_category)),
         pattern_class = dplyr::recode(pattern_class, pos = "pos_patterns", neg = "neg_patterns"))

motifs_compiled <- motifs_compiled_all %>% 
  # filter out exclude motifs
  filter(category != "exclude") %>% 
  # extract the name of the known motif
  separate(match0, into = c("TF0", "motif0"), sep = ": ")

# for cases with multiple patterns merged into one motif, take the one with the most
# number of seqlets.
motifs_compiled_unique <- motifs_compiled %>% 
  group_by(idx_uniq) %>% 
  slice_max(n = 1, order_by = tibble(total_n_seqlets, n_component_celltypes),
            with_ties = FALSE)

dim(motifs_compiled_unique)

# get positive and negative patterns
pos_patterns <- motifs_compiled_unique %>% filter(pattern_class == "pos_patterns") %>% pull(motif_name)
length(pos_patterns)

neg_patterns <- motifs_compiled_unique %>% filter(pattern_class == "neg_patterns") %>% pull(motif_name)
length(neg_patterns)

# initial breakdown of motif categories, prior manual collapsing
table(motifs_compiled_all$category)

# breakdown after manual collapsing & removing low-quality motifs
table(motifs_compiled_unique$category)

dim(motifs_compiled_unique)

```

Thus, in total we have:  

- **`r nrow(motifs_compiled_unique)` unique motifs** learned from the dataset 
- **`r length(pos_patterns)` positive patterns**
- **`r length(neg_patterns)` negative patterns**

_**NOTE:**_ There are 509 unique motifs, and the labels go from 0 to 508 (which spans 509 values). The 
"missing label" is for `idx_uniq` 466, which corresponded to unresolved motifs which
were filtered out.



Load reconciled hit calls:

```{r load_hits}

finemo_param <- "counts_v0.23_a0.8_all"

# load hit counts per motif per cell type
hits <- map(chrombpnet_models_keep$Cluster,
            ~ data.table::fread(here(glue("output/03-chrombpnet/02-compendium/hits_unified_motifs/reconciled_per_celltype_peaks/{.x}/{finemo_param}/hits_per_motif.tsv")),
                                data.table = FALSE) %>%
              dplyr::rename(n = count) %>% 
              tibble::add_column(Cluster = .x, .before = 1)) %>%
  setNames(chrombpnet_models_keep$Cluster)

# sanity check
n_cols <- map_dbl(hits, ncol)

if (any(n_cols < 12)) {
  
  missing_cols <- names(n_cols[n_cols < 12])
  stop("@ Missing columns in ", glue_collapse(missing_cols, ", "))
  
}

saveRDS(hits, file = glue("{out}/hits.Rds"))

# load summary stats on hits per peak per cell type
hits_per_peak <- map_dfr(chrombpnet_models_keep$Cluster,
                         ~ data.table::fread(here(glue("output/03-chrombpnet/02-compendium/hits_unified_motifs/reconciled_per_celltype_peaks/{.x}/{finemo_param}/hits_per_peak.tsv")),
                                             data.table = FALSE) %>%
                           tibble::add_column(Cluster = .x, .before = 1))

```



Calculate total number of hits per pattern per cell type:

```{r total_hits_per_pattern}

hits_all <- bind_rows(hits)

if (!all(motifs_compiled_unique$motif_name %in% hits_all$motif_name) | !all(hits_all$motif_name %in% motifs_compiled$motif_name)) {
  
  stop("@ Motif names in annotation don't match motif names in hits.")
  
}

hits_all_totals <- hits_all %>%
  group_by(motif_name) %>%
  summarize(total_hits = sum(n))

# organ w/ most hits
hits_top_organ <- hits_all %>% 
  left_join(cluster_meta %>% dplyr::select(Cluster = Cluster_ChromBPNet, organ)) %>% 
  group_by(motif_name, organ) %>% 
  summarize(hits_per_organ = sum(n)) %>% 
  group_by(motif_name) %>% 
  slice_max(n = 1, order_by = hits_per_organ, with_ties = FALSE) %>% 
  dplyr::select(motif_name, top_organ = organ)

dim(hits_all_totals)
dim(hits_top_organ)

motifs_compiled_unique <- hits_all_totals %>% 
  left_join(motifs_compiled_unique, by = "motif_name") %>% 
  left_join(hits_top_organ, by = "motif_name") %>% 
  mutate(motif_name_safe = gsub("\\/|\\|", ".", motif_name)) %>% 
  mutate(modisco_cwm_fwd = here(glue("output/03-chrombpnet/02-compendium/modisco_compiled/trimmed_logos/{pattern_class}.{pattern}.cwm.fwd.png")),
         modisco_cwm_rev = here(glue("output/03-chrombpnet/02-compendium/modisco_compiled/trimmed_logos/{pattern_class}.{pattern}.cwm.rev.png")))

# also subset to positive and neg patterns
motifs_compiled_unique_pos <- motifs_compiled_unique %>% 
  filter(pattern_class == "pos_patterns")

motifs_compiled_unique_neg <- motifs_compiled_unique %>% 
  filter(pattern_class == "neg_patterns")

dim(motifs_compiled_unique_pos)

hits_all_anno <- hits_all %>% 
  left_join(cluster_meta %>% dplyr::rename(Cluster_short = Cluster), by = c("Cluster" = "Cluster_ChromBPNet")) %>% 
  left_join(motifs_compiled_unique %>% dplyr::select(motif_name, pattern, annotation_broad, category))

head(hits_all_anno)

# breakdown of motif categories per pos patterns
table(motifs_compiled_unique_pos$category)

# save as TSV for re-use
motifs_compiled_unique %>% write_tsv(glue("{out}/motifs_compiled_unique.tsv"))

# short version
motifs_compiled_unique %>%
  arrange(idx_uniq) %>% 
  dplyr::select(motif_name, total_hits, idx, idx_uniq, pattern_class, pattern,
                annotation, annotation_broad, category,
                TF0, motif0, qval0, top_organ, component_organs, component_celltypes) %>% 
  write_tsv(glue("{out}/motifs_compiled_unique_short.tsv"))

hits_all_anno %>% write_tsv(glue("{out}/hits_per_cluster_annotated.tsv"))

```


Prep the hit counts matrices across cell types:

```{r prep_hits_mat}

dim(hits_all)

mat_hits <- hits_all %>%
  filter(pattern_class == "pos_patterns") %>% 
  left_join(motifs_compiled_unique_pos, by = c("motif_name", "pattern_class")) %>%
  filter(category %in% c("base", "base_with_flank", "homocomposite", "heterocomposite")) %>%
  dplyr::select(Cluster, motif_name, n) %>%
  spread(motif_name, n, fill = 0) %>%
  tibble::column_to_rownames(var = "Cluster") %T>%
  write_tsv(glue("{out}/hit_matrix.pos_patterns_no_unresolved.tsv"))

dim(mat_hits)

hits_all_broad <- hits_all %>%
  filter(pattern_class == "pos_patterns") %>% 
  left_join(motifs_compiled_unique_pos, by = "motif_name") %>%
  filter(category %in% c("base", "base_with_flank", "homocomposite", "heterocomposite")) %>%
  group_by(annotation_broad, Cluster) %>%
  summarize(n = sum(n))

mat_hits_broad <- hits_all_broad %>%
  spread(annotation_broad, n, fill = 0) %>%
  tibble::column_to_rownames(var = "Cluster") %T>%
  write_tsv(glue("{out}/hit_matrix.broad_pos_patterns_no_unresolved.tsv"))
dim(mat_hits_broad)

mat_hits_unresolved <- hits_all %>%
  filter(pattern_class == "pos_patterns") %>% 
  left_join(motifs_compiled_unique_pos, by = "motif_name") %>%
  filter(category == "unresolved") %>%
  dplyr::select(motif_name, n, Cluster) %>%
  spread(motif_name, n, fill = 0) %>%
  as.data.frame() %>%
  tibble::column_to_rownames(var = "Cluster")

dim(mat_hits_unresolved)

```


Load hit counts for binned distances to nucleosomes:

```{r load_motif_dist_dyad}

# load motif hit counts per binned distance to nucleosome dyads
hit_dyad_dist <- map_dfr(chrombpnet_models_keep$Cluster,
                         ~ data.table::fread(
                           here(glue("output/03-chrombpnet/02-compendium/nucleoatac/{.x}/{.x}.motif_dist_to_dyad.tsv")),
                           data.table = FALSE) %>%
                           tibble::add_column(Cluster = .x, .before = 1))

hit_dyad_dist <- left_join(hit_dyad_dist, motifs_compiled_unique, by = "motif_name")

```


Load hit counts for binned distances to peak summits:

```{r load_motif_dist_summit}

# load motif hit counts per binned distance to nucleosome dyads
hit_summit_dist <- map_dfr(chrombpnet_models_keep$Cluster,
                           ~ data.table::fread(
                             here(glue("output/03-chrombpnet/02-compendium/hits_unified_motifs/reconciled_per_celltype_peaks/{.x}/counts_v0.23_a0.8_all/motif_dist_to_summit.tsv")),
                             data.table = FALSE) %>%
                             tibble::add_column(Cluster = .x, .before = 1))

hit_summit_dist <- left_join(hit_summit_dist, motifs_compiled_unique, by = "motif_name")

```


We also count the number of unique peaks in each cluster that contain at least one motif instance,
separately for positive and negative patterns:

```{bash, eval = FALSE}

cd ../../../output/03-chrombpnet/02-compendium/hits_unified_motifs/reconciled_per_celltype_peaks
for class in ( pos_patterns neg_patterns ); do
for i in *; do
n_peaks=`zcat ${i}/counts_v0.23_a0.8_all/hits_unique.reconciled.annotated.tsv.gz | grep ${class} | cut -f 14 | sort | uniq | wc -l`
echo -e "${i}\t${n_peaks}" >> .unique_peaks_cluster.${class}.txt
done
done

```


Load the counts:

```{r n_peaks}

peaks_w_pos_hits <- read_tsv(here("output/03-chrombpnet/02-compendium/hits_unified_motifs/reconciled_per_celltype_peaks/.unique_peaks_cluster.pos_patterns.txt"),
                             col_names = c("Cluster", "N_peaks_w_hits")) %>% 
  mutate(pattern_class = "pos_patterns")

peaks_w_neg_hits <- read_tsv(here("output/03-chrombpnet/02-compendium/hits_unified_motifs/reconciled_per_celltype_peaks/.unique_peaks_cluster.neg_patterns.txt"),
                             col_names = c("Cluster", "N_peaks_w_hits")) %>% 
  mutate(pattern_class = "neg_patterns")

peaks_w_hits <- bind_rows(peaks_w_pos_hits,
                          peaks_w_neg_hits) %>%
  left_join(cluster_meta %>% dplyr::select(organ, Cluster = Cluster_ChromBPNet, n_peaks), by = "Cluster") %>% 
  mutate(Prop_peaks_w_hits = N_peaks_w_hits / n_peaks)

```


Load HOCOMOCO:

```{r hocomoco}

hocomoco <- read_tsv(here("output/02-global_analysis/04/hocomoco_unique.tsv"))

unique_classes <- unique(hocomoco$masterlist_info.tfclass_class)
cmap_class <- c(ArchR::paletteDiscrete(unique_classes, set = "stallion"), "N/A" = "gray90")

```



Load CWMs for the compiled motifs:

```{r known_cwm}

# use custom helper
cwm_list <- read_cwm_meme(here(glue("output/03-chrombpnet/02-compendium/modisco_compiled/modisco_compiled.memedb.txt")))
length(cwm_list)

cwm_list <- cwm_list[base::intersect(names(cwm_list), motifs_compiled_unique$pattern)]
length(cwm_list)

names(cwm_list) <- plyr::mapvalues(names(cwm_list), from = motifs_compiled_unique$pattern, to = motifs_compiled_unique$motif_name)

# label matrix dimensions
cwm_list <- map(cwm_list, function(cwm) {
  
  rownames(cwm) <- c("A", "C", "G", "T")
  colnames(cwm) <- seq(1:30)
  
  return(cwm)
  
})

print(length(cwm_list))
saveRDS(cwm_list, file = glue("{out}/denovo_cwm_list.Rds"))

```


## Save intermediate objects

```{r}

save(chrombpnet_models_keep, cluster_meta, motifs_compiled_unique, motifs_compiled_all,
     cwm_list, hit_dyad_dist, hit_summit_dist, hits_all_anno, hits_per_peak,
     file = glue("{out}/intermediate_obj.Rda"))

```


To reload for later analysis:

```{r, eval = FALSE}

load(glue("{out}/intermediate_obj.Rda"))
hits <- readRDS(glue("{out}/hits.Rds"))
hits_all <- bind_rows(hits)

```


## Rename trimmed CWM logos

```{r, eval = FALSE}

fs::dir_create(file.path(out, "trimmed_cwm_logos"))

all(fs::file_exists(motifs_compiled_unique$modisco_cwm_fwd))
all(fs::file_exists(motifs_compiled_unique$modisco_cwm_rev))

for (i in 1:nrow(motifs_compiled_unique)) {
  
  new_name_fw <- paste0(motifs_compiled_unique$motif_name_safe[i], "__cwm.fwd.png")
  fs::file_copy(motifs_compiled_unique$modisco_cwm_fwd[i], file.path(out, "trimmed_cwm_logos", new_name_fw), overwrite = TRUE)
  
  new_name_rev <- paste0(motifs_compiled_unique$motif_name_safe[i], "__cwm.rev.png")
  fs::file_copy(motifs_compiled_unique$modisco_cwm_rev[i], file.path(out, "trimmed_cwm_logos", new_name_rev), overwrite = TRUE)
  
}

```



## Prep table

```{r}

motifs_compiled_unique %>% 
  mutate(
    class = case_when(
      pattern_class == "pos_patterns" ~ "positive",
      TRUE ~ "negative"
    ),
    cwm_logo_fwd = paste0(motif_name_safe, "__cwm.fwd.png"),
    cwm_logo_rev = paste0(motif_name_safe, "__cwm.rev.png"),
  ) %>% 
  dplyr::select(
    idx_uniq,
    motif_name,
    motif_name_safe,
    class,
    total_instances_across_celltypes = total_hits,
    cwm_logo_fwd,
    cwm_logo_rev,
    query_consensus,
    annotation,
    annotation_broad,
    category,
    best_match_TF = TF0,
    best_match_motif = motif0,
    best_match_TOMTOM_qval = qval0,
    best_match_TFs_in_family = match0_TFs_in_family,
    best_match_TFs_in_Vierstra_archetype = match0_TFs_in_archetype,
    total_seqlets_across_celltypes = total_n_seqlets,
    merged_pattern = pattern,
    n_constituent_motifs = n_component_patterns,
    n_constituent_celltypes = n_component_celltypes,
    constituent_organs = component_organs,
    constituent_celltypes = component_celltypes
  ) %>% 
  arrange(idx_uniq) %>% 
  write_csv(glue("{out}/TABLE_motif_compendium.csv"))

```


# Global view

## Categories of motifs

What kinds of patterns did we detect?

First, let's look at all 834 motifs from the motif clustering workflow.

```{r breakdown_category, fig.width = 11, fig.height = 6}

p1 <- motifs_compiled_all %>% 
  group_by(category) %>% 
  count() %>% 
  # reverse levels since we're flipping coordinates
  mutate(category = factor(category, levels = rev(names(cmap_category)))) %>% 
  ggplot(aes(x = category, y = n)) +
  geom_col(aes(fill = category)) +
  geom_text(aes(label = n), hjust = -0.5, fontface = "bold", size = 4) +
  coord_flip() +
  scale_fill_manual(values = cmap_category) +
  ggtitle("De novo deep learning derived motifs \n(after MoDISco agggregation step)") +
  no_legend() +
  theme(panel.grid.major.x = element_line(color = "gray90"),
        panel.grid.minor.x = element_line(color = "gray90")) +
  ylim(c(0, 280))

p2 <- motifs_compiled_all %>% 
  ggplot(aes(x = "breakdown")) +
  geom_bar(aes(fill = category), position = "fill") +
  scale_fill_manual(values = cmap_category)

plot_grid(p1, p2, nrow = 1, align = "h", axis = "tb", rel_widths = c(0.65, 0.35))

```


Next, let's look at the unique motifs after QC and collapsing motifs:

```{r barplot_n_motifs, fig.width = 10, fig.height = 5}

# stacked barplot
p1 <- motifs_compiled_unique %>% 
  ggplot(aes(x = "breakdown")) +
  geom_bar(aes(fill = category), position = "fill") +
  scale_fill_manual(values = cmap_category) +
  xlab(NULL) +
  theme(legend.position = "bottom") +
  no_legend()

# barplot of counts per category
p2 <- motifs_compiled_unique %>% 
  group_by(category) %>% 
  count() %>% 
  # reverse levels since we're flipping coordinates
  mutate(category = factor(category, levels = rev(names(cmap_category)))) %>% 
  ggplot(aes(x = category, y = n)) +
  geom_col(aes(fill = category)) +
  geom_text(aes(label = n), hjust = -0.5, fontface = "bold", size = 4) +
  coord_flip() +
  scale_fill_manual(values = cmap_category) +
  ggtitle("De novo deep learning derived motifs \n(after manual collapsing & removing low quality motifs)") +
  no_legend() +
  theme(panel.grid.major.x = element_line(color = "gray90"),
        panel.grid.minor.x = element_line(color = "gray90")) +
  ylim(c(0, 280))


plot_grid(p1, p2, nrow = 1, align = "h", axis = "tb", rel_widths = c(1, 4, 3.5))

```


## Positive and negative patterns

```{r pos_neg_breakdown, fig.width = 7, fig.height = 9}

p1 <- motifs_compiled_unique %>% 
  group_by(pattern_class) %>% 
  count() %>% 
  ggplot(aes(x = "1", y = n)) +
  geom_col(aes(fill = pattern_class), alpha = 0.4) +
  coord_flip() +
  scale_fill_manual(values = c("pos_patterns" = "darkgreen", "neg_patterns" = "red")) +
  ggtitle("De novo deep learning derived motifs") +
  no_legend() +
  theme(panel.grid.major.x = element_line(color = "gray90"),
        panel.grid.minor.x = element_line(color = "gray90"))

p2 <- hits_all %>%
  left_join(cluster_meta %>% dplyr::select(Cluster = Cluster_ChromBPNet, organ)) %>% 
  group_by(organ, pattern_class) %>%
  summarize(n_hits = sum(n)) %>%
  mutate(organ = factor(organ, levels = rev(unique(.$organ)))) %>% 
  ggplot(aes(x = organ, y = n_hits)) +
  geom_col(aes(fill = pattern_class), alpha = 0.4, position = "fill") +
  coord_flip() +
  scale_fill_manual(values = c("pos_patterns" = "darkgreen", "neg_patterns" = "red")) +
  ggtitle("Breakdown of total hits") +
  no_legend() +
  theme(panel.grid.major.x = element_line(color = "gray90"),
        panel.grid.minor.x = element_line(color = "gray90"))

plot_grid(p1, p2, ncol = 1, align = "v", axis = "rl", rel_heights = c(1, 4))

```


## Similarity to known motifs

Distribution of TOMTOM q-values for similarity to known motifs, and total # hits per motif:

```{r boxplot_nhits_and_similarity, fig.width = 10, fig.height = 5}

p1 <- motifs_compiled_unique %>%
  mutate(similarity = qval0) %>%
  ggplot(aes(x = category, y = -log10(qval0))) +
  geom_boxplot(aes(fill = category), outliers = FALSE) +
  geom_jitter(shape = 21, width = 0.2, size = 0.5) +
  scale_fill_manual(values = cmap_category) +
  geom_hline(yintercept = -log10(0.05)) +
  no_legend() +
  xlab(NULL) +
  rotate_x()

p2 <- motifs_compiled_unique %>% 
  ggplot(aes(x = category, y = log10(total_hits))) +
  geom_boxplot(aes(fill = category), outlier.colour = NA) +
  geom_jitter(color = "black", width = 0.2, size = 0.5, shape = 21) +
  scale_fill_manual(values = cmap_category) +
  rotate_x() +
  no_legend() +
  # ggtitle("Total # of hits per unique motif \neach point is a motif (n=525)") +
  coord_cartesian(ylim = c(1, 7))

plot_grid(p1, p2, nrow = 1, align = "h", axis = "tb")

```


Total number of hits per cell type:

```{r total_n_hits, fig.width = 22, fig.height = 8}

hits_all %>%
  group_by(Cluster, pattern_class) %>%
  summarize(total_hits = sum(n)) %>% 
  left_join(cluster_meta, by = c("Cluster" = "Cluster_ChromBPNet")) %>% 
  arrange(organ, L1_annot) %>% 
  mutate(Cluster = factor(Cluster, levels = unique(.$Cluster))) %>% 
  ggplot(aes(x = Cluster, y = total_hits)) +
  facet_wrap(~ pattern_class, ncol = 1) +
  geom_bar(stat = "identity", aes(fill = organ)) +
  scale_fill_manual(values = cmap_organ) +
  rotate_x() +
  ylab("Total number of hits") +
  theme(
    strip.background = element_blank(),
    axis.text.x = element_text(size = 5),
    panel.grid.major.y = element_line(color = "gray90"),
    panel.grid.minor.y = element_line(color = "gray90")) +
  scale_y_continuous(labels = scales::comma) +
  ggtitle("Total hits per cell type")

```



## Total hits

By class (positive/negative patterns):

```{r total_n_hits_by_class, fig.width = 18, fig.height = 6}

hits_all %>%
  left_join(motifs_compiled_unique %>% dplyr::select(-pattern_class),
            by = c("motif_name")) %>%
  group_by(Cluster, pattern_class) %>%
  summarize(n_per_class = sum(n)) %>% 
  left_join(cluster_meta, by = c("Cluster" = "Cluster_ChromBPNet")) %>% 
  arrange(organ, L1_annot) %>% 
  mutate(Cluster = factor(Cluster, levels = unique(.$Cluster))) %>% 
  ggplot(aes(x = Cluster, y = n_per_class)) +
  geom_bar(stat = "identity", aes(fill = pattern_class), alpha = 0.4, position = "fill") +
  scale_fill_manual(values = c("pos_patterns" = "darkgreen", "neg_patterns" = "red")) +
  rotate_x() +
  ylab("Total number of hits") +
  theme(
    axis.text.x = element_text(size = 5),
    panel.grid.major.y = element_line(color = "gray90"),
    panel.grid.minor.y = element_line(color = "gray90")) +
  scale_y_continuous(labels = scales::comma) +
  ggtitle("Total hits, by pattern class")

```


Let's plot total hits and mean hits per peak, grouped by organ:


```{r nhits_per_cluster_aggregated_plot, fig.width = 10, fig.height = 6}

peaks_w_hits %>% 
  group_by(pattern_class, organ) %>% 
  summarize(median = median(Prop_peaks_w_hits)) %>% 
  spread(pattern_class, median)

p0 <- peaks_w_hits %>% 
  mutate(pattern_class = factor(pattern_class, levels = c("pos_patterns", "neg_patterns"))) %>% 
  ggplot(aes(x = organ, y = Prop_peaks_w_hits)) +
  geom_jitter(stat = "identity", aes(fill = organ), shape = 21, color = "white", size = 2, alpha = 0.6, width = 0.25) +
  scale_fill_manual(values = cmap_organ) +
  rotate_x() +
  facet_wrap(~ pattern_class, ncol = 1) +
  ylab("Proportion") +
  theme(
    strip.background = element_blank(),
    panel.grid.major.y = element_line(color = "gray90"),
    panel.grid.minor.y = element_line(color = "gray90")) +
  scale_y_continuous(labels = scales::comma) +
  ggtitle("Peaks with hits") +
  no_legend()

p1 <- hits_all %>% 
  group_by(Cluster, pattern_class) %>%
  summarize(total_hits = sum(n)) %>% 
  left_join(cluster_meta, by = c("Cluster" = "Cluster_ChromBPNet")) %>% 
  arrange(organ, L1_annot) %>% 
  mutate(Cluster = factor(Cluster, levels = unique(.$Cluster)),
         pattern_class = factor(pattern_class, levels = c("pos_patterns", "neg_patterns"))) %>% 
  ggplot(aes(x = organ, y = total_hits)) +
  geom_jitter(stat = "identity", aes(fill = organ), shape = 21, color = "white", size = 2, alpha = 0.6, width = 0.25) +
  scale_fill_manual(values = cmap_organ) +
  rotate_x() +
  facet_wrap(~ pattern_class, ncol = 1) +
  ylab("Total number of hits") +
  theme(
    strip.background = element_blank(),
    panel.grid.major.y = element_line(color = "gray90"),
    panel.grid.minor.y = element_line(color = "gray90")) +
  scale_y_continuous(labels = scales::comma) +
  ggtitle("Total hits") +
  no_legend()

p2 <- hits_per_peak %>% 
  left_join(cluster_meta, by = c("Cluster" = "Cluster_ChromBPNet")) %>% 
  arrange(organ, L1_annot) %>% 
  mutate(Cluster = factor(Cluster, levels = unique(.$Cluster))) %>% 
  gather(pattern_class, median_n_hits, median_pos_patterns, median_neg_patterns) %>% 
  mutate(pattern_class = factor(pattern_class, levels = c("median_pos_patterns", "median_neg_patterns"))) %>% 
  ggplot(aes(x = organ, y = median_n_hits)) +
  ggbeeswarm::geom_quasirandom(
    aes(fill = organ), shape = 21, color = "white", size = 2, alpha = 0.6) + #, width = 0.25, height = 0) +
  scale_fill_manual(values = cmap_organ) +
  facet_wrap(~ pattern_class, ncol = 1) +
  rotate_x() +
  ylab("Median number of hits per peak") +
  theme(
    strip.background = element_blank(),
    panel.grid.major.y = element_line(color = "gray90"),
    panel.grid.minor.y = element_line(color = "gray90")) +
  scale_y_continuous(breaks = seq(0, 10, 1), labels = seq(0, 10), limits = c(0, 9)) +
  no_legend() +
  ggtitle("Median hits per peak")

plot_grid(p0, p1, p2, nrow = 1, align = "h", axis = "tb", rel_widths = c(1, 1.2, 1))

```


## Effects of coverage/peaks on number of hits

Here we look at how total number of motif instances varies with total fragments
and number of peaks:

```{r scatter_peaks_vs_frags_vs_instances, fig.width = 15, fig.height = 8}

# calculate total hist per cluster (incl. both pos/neg patterns)
total_hits_per_cluster <- hits_all %>% 
  group_by(Cluster) %>%
  summarize(total_hits = sum(n)) %>% 
  left_join(cluster_meta, by = c("Cluster" = "Cluster_ChromBPNet"))

# this is for the 189 models which passed QC
total_hits_per_cluster$Cluster %>% unique %>% length

summary(total_hits_per_cluster$total_hits)

cor_frags <- lm(total_hits_per_cluster$total_hits ~ total_hits_per_cluster$total_frags)
summary(cor_frags)

cor_peaks <- lm(total_hits_per_cluster$total_hits ~ total_hits_per_cluster$n_peaks)
summary(cor_peaks)

p1 <- total_hits_per_cluster %>% 
  ggplot(aes(x = log10(total_frags), y = total_hits)) +
  geom_point(aes(fill = organ), shape = 21, size = 3, alpha = 0.7, color = "white") +
  scale_fill_manual(values = cmap_organ) +
  scale_y_continuous(labels = label_number(scale = 1/1e6)) +
  annotate("text", label = glue("cor = {round(sqrt(summary(cor_frags)$r.squared), 3)}, \nR^2 = {round(summary(cor_frags)$r.squared, 3)}"),
           x = 7, y = 2e6, size = 5) +
  xlab("log10(Total fragments)") + ylab("Total hits (x 1M)") +
  square() +
  ggtitle("# motif instances vs. \n# fragments") +
  no_legend()

p2 <- total_hits_per_cluster %>% 
  ggplot(aes(x = n_peaks, y = total_hits)) +
  geom_point(aes(fill = organ), shape = 21, size = 3, alpha = 0.7, color = "white") +
  scale_fill_manual(values = cmap_organ) +
  scale_y_continuous(labels = label_number(scale = 1/1e6)) +
  scale_x_continuous(labels = scales::comma) +
  annotate("text", label = glue("cor = {round(sqrt(summary(cor_peaks)$r.squared), 3)}, \nR^2 = {round(summary(cor_peaks)$r.squared, 3)}"),
           x = 7e4, y = 2e6, size = 5) +
  xlab("Total peaks") + ylab("Total hits (x 1M)") +
  square() +
  ggtitle("# motif instances vs. \n# peaks") +
  no_legend()

# median number of instances for positive motifs per peaks
n_hits_per_peaks <- hits_per_peak %>% 
  left_join(cluster_meta, by = c("Cluster" = "Cluster_ChromBPNet"))

cor_frags <- lm(n_hits_per_peaks$median_pos_patterns ~ n_hits_per_peaks$total_frags)
summary(cor_frags)

cor_peaks <- lm(n_hits_per_peaks$median_pos_pattern ~ n_hits_per_peaks$n_peaks)
summary(cor_peaks)

p3 <- n_hits_per_peaks %>% 
  ggplot(aes(x = log10(total_frags), y = median_pos_patterns)) +
  geom_point(aes(fill = organ), shape = 21, size = 3, alpha = 0.7, color = "white") +
  scale_fill_manual(values = cmap_organ) +
  scale_y_continuous(labels = scales::comma) +
  no_legend() +
  square() +
  annotate("text", label = glue("cor = {round(sqrt(summary(cor_frags)$r.squared), 3)}, \nR^2 = {round(summary(cor_frags)$r.squared, 3)}"),
           x = 7.5, y = 7, size = 5) +
  xlab("log10(Total fragments)") + ylab("median # positive hits per peak") +
  ggtitle("median positive hits per peak vs. \n# fragments")

p4 <- n_hits_per_peaks %>% 
  ggplot(aes(x = n_peaks, y = median_pos_patterns)) +
  geom_point(aes(fill = organ), shape = 21, size = 3, alpha = 0.7, color = "white") +
  scale_fill_manual(values = cmap_organ) +
  scale_y_continuous(labels = scales::comma) + 
  scale_x_continuous(labels = scales::comma) +
  no_legend() +
  square() +
  annotate("text", label = glue("cor = {round(sqrt(summary(cor_peaks)$r.squared), 3)}, \nR^2 = {round(summary(cor_peaks)$r.squared, 3)}"),
           x = 70000, y = 7, size = 5) +
  xlab("total number of peaks") + ylab("median # positive hits per peak") +
  ggtitle("median positive hits per peak vs. \n# peaks")

# calculate quartiles
n_hits_per_peaks$n_peak_quartile <- ntile(n_hits_per_peaks$n_peaks, n=4)
table(n_hits_per_peaks$n_peak_quartile)/sum(table(n_hits_per_peaks$n_peak_quartile))

n_hits_per_peaks$total_frags_quartile <- ntile(n_hits_per_peaks$total_frags, n=4)
table(n_hits_per_peaks$total_frags_quartile)/sum(table(n_hits_per_peaks$total_frags_quartile))

p5 <- n_hits_per_peaks %>% 
  ggplot(aes(x = total_frags_quartile, y = median_pos_patterns, group = total_frags_quartile)) +
  geom_boxplot(fill = "gray90", outliers = FALSE) +
  ggbeeswarm::geom_quasirandom(
    aes(fill = organ), shape = 21, color = "white", size = 4, alpha = 0.6, width = 0.2) +
  scale_fill_manual(values = cmap_organ) +
  no_legend() +
  ylab("Median # positive hits per peak") +
  xlab("Quantile (total number of fragments)") +
  ggtitle("median positive hits per peak vs. \n# fragments")

p6 <- n_hits_per_peaks %>% 
  ggplot(aes(x = n_peak_quartile, y = median_pos_patterns, group = n_peak_quartile)) +
  geom_boxplot(fill = "gray90", outliers = FALSE) +
  ggbeeswarm::geom_quasirandom(
    aes(fill = organ), shape = 21, color = "white", size = 4, alpha = 0.6, width = 0.2) +
  scale_fill_manual(values = cmap_organ) +
  no_legend() +
  ylab("Median # positive hits per peak") +
  xlab("Quantile (total number of peaks)") +
  ggtitle("median positive hits per peak vs. \n# peaks")

plot_grid(p1, p3, p5, p2, p4, p6, align = "hv", axis = "rltb")

```


# Genomic localization

## Nucleosome positioning

First we need to aggregate the counts per bin over motif variants and cell types,
per category:

```{r aggregate_nuc_pos}

# total counts
motifs_keep <- hit_dyad_dist %>% 
  filter(category %in% c("base", "base_with_flank")) %>% 
  group_by(annotation_broad) %>% 
  summarize(sum_all_counts = sum(n)) %>%
  arrange(sum_all_counts) %>% 
  filter(sum_all_counts > 50000) %>%
  pull(annotation_broad)

motifs_keep %>% length

medians_dist_dyad <- hit_dyad_dist %>% 
  filter(category %in% c("base", "base_with_flank") &
           annotation_broad %in% motifs_keep) %>% 
  group_by(annotation_broad, bin_dist) %>% 
  summarize(sum_counts = sum(n)) %>% 
  # we need to actually calculate the median given the binned counts
  slice_max(order_by = sum_counts, n = 1) %>% 
  arrange(bin_dist)

# z-score
hit_dyad_dist_summed <- hit_dyad_dist %>% 
  filter(category %in% c("base", "base_with_flank") &
           annotation_broad %in% motifs_keep) %>% 
  group_by(annotation_broad, bin_dist) %>% 
  summarize(sum_counts = sum(n)) %>% 
  # zscore counts per motif (row-wise)
  mutate(zscore = scale(sum_counts)) %>% 
  mutate(annotation_broad = factor(annotation_broad, levels = unique(medians_dist_dyad$annotation_broad)),
         bin_dist = as.numeric(bin_dist)) %>% 
  ungroup()

```

```{r nuc_heatmap}

p1 <- hit_dyad_dist_summed %>%
  ggplot(aes(x = bin_dist, y = annotation_broad)) +
  geom_tile(aes(fill = zscore)) +
  scale_fill_gradientn(colours = rdbu2,
                       # set midpoint to 0
                       # https://github.com/tidyverse/ggplot2/issues/3738#issuecomment-1336166757
                       rescaler = ~ scales::rescale_mid(.x, mid = 0), limits = c(-3, 4)) +
  geom_vline(xintercept = 75, linewidth = 1) +
  scale_x_continuous(breaks = seq(10, 250, by = 10)) +
  theme(panel.border = element_blank(),
        axis.ticks = element_blank(),
        axis.text.x = element_text(size = 5),
        axis.text.y = element_text(size = 9)) +
  xlab("binned distance from nucleosome dyad (bp)") +
  ylab("motif") +
  ggtitle("Motif instances distances to nucleosome dyad \n(aggregated within categories, across cell types)")

```



We can plot the heatmap w/ the actual counts on top so that we can manually inspect:

```{r nucleosome_positioning_heatmap, fig.width = 14, fig.height = 8}

hit_dyad_dist_summed %>%
  ggplot(aes(x = bin_dist, y = annotation_broad)) +
  geom_tile(aes(fill = zscore)) +
  geom_text(aes(label = round(sum_counts, 2)), color = "black") +
  scale_fill_gradientn(colours = rdbu2,
                       # set midpoint to 0
                       # https://github.com/tidyverse/ggplot2/issues/3738#issuecomment-1336166757
                       rescaler = ~ scales::rescale_mid(.x, mid = 0)) +
  geom_vline(xintercept = 75, linewidth = 1) +
  scale_x_continuous(breaks = seq(10, 250, by = 10)) +
  theme(panel.border = element_blank(),
        axis.ticks = element_blank(),
        axis.text.x = element_text(size = 5),
        axis.text.y = element_text(size = 9)) +
  xlab("binned distance from nucleosome dyad (bp)") +
  ylab("motif") +
  ggtitle("Motif instances distances to nucleosome dyad \n(aggregated within categories, across cell types)")

```



## Distance to summits


```{r peak_flank_heatmap}

hit_summit_dist_summed <- hit_summit_dist %>% 
  filter(category %in% c("base", "base_with_flank") &
           annotation_broad %in% motifs_keep) %>% 
  group_by(annotation_broad, bin_dist) %>% 
  summarize(sum_counts = sum(n)) %>% 
  filter(bin_dist <= 250) %>% 
  # zscore counts per motif (row-wise)
  mutate(zscore = scale(sum_counts)) %>% 
  mutate(annotation_broad = factor(annotation_broad, levels = unique(medians_dist_dyad$annotation_broad)),
         bin_dist = as.numeric(bin_dist)) %>% 
  ungroup()

hit_summit_dist_summed %>%
  ggplot(aes(x = bin_dist, y = annotation_broad)) +
  geom_tile(aes(fill = sum_counts)) +
  scale_fill_viridis_c("plasma") +
  scale_x_continuous(breaks = seq(0, 250, by = 10)) +
  theme(panel.border = element_blank(),
        axis.ticks = element_blank(),
        axis.text.x = element_text(size = 5),
        axis.text.y = element_text(size = 9)) +
  xlab("binned distance from peak summit (bp)") +
  ylab("motif") +
  ggtitle("Motif instances distances to peak summit \n(aggregated within categories, across cell types)")

p2 <- hit_summit_dist_summed %>%
  ggplot(aes(x = bin_dist, y = annotation_broad)) +
  geom_tile(aes(fill = zscore)) +
  scale_fill_gradientn(colours = rdbu2,
                       # set midpoint to 0
                       # https://github.com/tidyverse/ggplot2/issues/3738#issuecomment-1336166757
                       rescaler = ~ scales::rescale_mid(.x, mid = 0), limits = c(-3, 4)) +
  scale_x_continuous(breaks = seq(0, 250, by = 10)) +
  theme(panel.border = element_blank(),
        axis.ticks = element_blank(),
        axis.text.x = element_text(size = 5),
        axis.text.y = element_text(size = 9)) +
  xlab("binned distance from peak summit (bp)") +
  ylab("motif") +
  ggtitle("Motif instances distances to peak summit \n(aggregated within categories, across cell types)")

```


```{r cdfs_peak_summits, fig.width = 5, fig.height = 6}

hit_summit_dist_summed <- hit_summit_dist %>% 
  filter(category %in% c("base", "base_with_flank") &
           annotation_broad %in% motifs_keep) %>% 
  group_by(annotation_broad, bin_dist, pattern_class) %>% 
  summarize(sum_counts = sum(n))

hit_summit_dist_summed %>%
  filter(annotation_broad != "unresolved") %>% 
  group_by(annotation_broad) %>%
  mutate(cum_sum_counts = cumsum(sum_counts)) %>%
  mutate(pattern_class = factor(pattern_class, levels = c("pos_patterns", "neg_patterns"))) %>% 
  ggplot(aes(x = bin_dist, y = log10(cum_sum_counts), group = annotation_broad)) +
  # geom_point(aes(fill = pattern_class), alpha = 0.4, color = "white", shape = 21) +
  geom_line(aes(color = pattern_class), alpha = 0.4) + 
  facet_wrap(~ pattern_class, ncol = 1) +
  scale_color_manual(values = c("pos_patterns" = "#BD202E", "neg_patterns" = "#1D75BC")) +
  scale_fill_manual(values = c("pos_patterns" = "#BD202E", "neg_patterns" = "#1D75BC")) +
  theme(panel.grid.minor.x = element_line(color = "gray90"),
        panel.grid.major.x = element_line(color = "gray90"),
        strip.background.x = element_blank()) +
  no_legend()

```


```{r barplot_peak_summits, fig.width = 11, fig.height = 5}

hit_summit_dist_coarse <- hit_summit_dist %>% 
  dplyr::select(Cluster, motif_name, annotation_broad, category, pattern_class, bin_dist, n_per_cluster_per_bin = n) %>% 
  filter(category %in% c("base", "base_with_flank") &
           annotation_broad %in% motifs_keep) %>% 
  mutate(bin_dist_coarse = case_when(
    bin_dist <= 50 ~ "0-50 bp",
    bin_dist <= 100 ~ "51-100 bp",
    bin_dist <= 150 ~ "101-150 bp",
    bin_dist <= 200 ~ "151-200 bp",
    bin_dist <= 250 ~ "201-250 bp",
    TRUE ~ "> 251 bp"
  )) %>% 
  group_by(annotation_broad, pattern_class, bin_dist_coarse) %>% 
  summarize(n_per_bin = sum(n_per_cluster_per_bin)) %>% 
  mutate(n_total = sum(n_per_bin),
         prop_ber_bin = n_per_bin / n_total)

cmap_bins <- RColorBrewer::brewer.pal(6, "BuGn")
names(cmap_bins) <- rev(c("0-50 bp", "51-100 bp", "101-150 bp", "151-200 bp", "201-250 bp", "> 251 bp"))

motif_order_dist <- hit_summit_dist_coarse %>% filter(bin_dist_coarse == "0-50 bp") %>% arrange(desc(prop_ber_bin)) %>% pull(annotation_broad) %>% unique()

hit_summit_dist_coarse %>% 
  mutate(bin_dist_coarse = factor(bin_dist_coarse, levels = names(cmap_bins)),
         annotation_broad = factor(annotation_broad, levels = motif_order_dist)) %>% 
  ggplot(aes(x = annotation_broad, y = prop_ber_bin)) +
  geom_bar(stat = "identity", aes(fill = bin_dist_coarse)) +
  scale_fill_manual(values = cmap_bins) +
  rotate_x()

```



Get proportions of motifs per bin:

```{r}

hit_summit_dist_coarse %>% 
  mutate(bin_dist_coarse = factor(bin_dist_coarse, levels = rev(names(cmap_bins)))) %>% 
  group_by(bin_dist_coarse) %>% 
  summarize(n_per_bin_across_motifs = sum(n_per_bin)) %>% 
  mutate(n_total = sum(n_per_bin_across_motifs),
         prop = n_per_bin_across_motifs / n_total,
         cum_n = cumsum(n_per_bin_across_motifs),
         cum_prop = cum_n / n_total) %>% 
  arrange(bin_dist_coarse)

```




## Feature type

```{r heatmap_genomic_localization, fig.width = 15, fig.height = 8}

region_type_summed <- hits_all_anno %>% 
  filter(annotation_broad %in% motifs_keep) %>% 
  dplyr::select(motif_name, annotation_broad, Distal, Exonic, Intronic, Promoter) %>% 
  gather(region_type, count, 3:ncol(.)) %>% 
  group_by(annotation_broad, region_type) %>% 
  summarize(count_total = sum(count, na.rm = TRUE)) %>% 
  mutate(annotation_broad = factor(annotation_broad, levels = unique(medians_dist_dyad$annotation_broad)))

p3 <- region_type_summed %>% 
  ggplot(aes(x = annotation_broad, y = count_total)) +
  geom_col(aes(fill = region_type), position = "fill", alpha = 0.7) +
  scale_fill_manual(values = cmap_peaktype2) +
  coord_flip() +
  theme(legend.position = "bottom")

plot_grid(p1, p2 , p3, nrow = 1, align = "h", axis = "tb", rel_widths = c(2, 2, 1))

```



# Summary plots of discovered motifs

Load known motifs from Vierstra input set:

```{r}

vierstra_motifs <- read_cwm_meme(file.path(kundaje_dir, "refs/Vierstra_motifs/all.dbs.meme"))

```




We want to describe motifs in a given subset. This plots:

- The _de novo_ CWM (trimmed)
- The PWM for the best matching known motif in the Vierstra input database
- Total # of hits across cell types
- Number of motif variants within a category
- Breakdown of hits by compartment
- Breakdown of hits by organ
- Genomic annotations
- Binned distances to nucleosomes


## Base motifs, representative per broad anno

```{r motif_summary_broad_base, fig.width = 24, fig.height = 14}

base_broad <- motifs_compiled_unique %>% 
  filter(category %in% c("base", "base_with_flank"))

length(base_broad$annotation_broad %>% unique())

plot_motif_summary(base_broad,
                   hits_all_anno,
                   hit_dyad_dist,
                   order_by = "distToTSS",
                   plot_known_motifs = TRUE,
                   # manually indicate which motifs to reverse complement
                   # to match known motif orientation
                   to_revcomp = c("GATA", "CTCF", "RFX", "MEF2", "NFKB/RELA", "NR", "ONECUT",
                                  "ZNF143", "NR:HNF4", "REST", "THRA/B", "HD:SIX/ZNF", "ZEB/SNAI", "YY1/2rep"))

```



## Homocomposites, representative motif per broad anno

```{r motifs_broad_homocomposites, fig.width = 24, fig.height = 9}

homocomp_broad <- motifs_compiled_unique %>%
  filter(pattern_class == "pos_patterns" & category == "homocomposite")

length(unique(homocomp_broad$annotation_broad))

plot_motif_summary(homocomp_broad,
                   hits_all_anno,
                   hit_dyad_dist,
                   plot_known_motifs = TRUE,
                   rel_widths = c(3.5, 2, 0.5, 1, 1, 1, 1, 1, 1, 1, 1),
                   to_revcomp = c("ETS_ETS", "IRF/STAT_IRF/STAT", "BZIP_BZIP", "HD:PITX/OTX_HD:PITX/OTX"))

```


## Heterocomposites, representative motif per broad anno

Just showing the top 30 here, because there are many.

```{r motifs_broad_heterocomposites, fig.width = 24, fig.height = 15}

heterocomp_broad <- motifs_compiled_unique %>% 
  filter(category == "heterocomposite" & !grepl("unresolved", annotation))

heterocomp_broad %>% 
  group_by(annotation_broad) %>% 
  count() %>% 
  arrange(desc(n))

plot_motif_summary(heterocomp_broad,
                   hits_all_anno,
                   hit_dyad_dist,
                   plot_known_motifs = FALSE,
                   # known_motifs = jolma_motifs,
                   # known_motif_col = "match0_jolma",
                   # known_motif_name = "match0_jolma",
                   rel_widths = c(3, 0.5, 1, 1, 1, 1, 1, 1.5, 1.5, 1),
                   top_n = 30)

```



## Negative/reducing patterns, representative per broad anno

```{r motif_summary_broad_neg, fig.width = 23, fig.height = 7}

neg_patterns <- motifs_compiled_unique %>% 
  filter(pattern_class == "neg_patterns") %>% 
  filter(!(category == "unresolved"))

# we can make use of the plotting function by setting the broad annotation to 
# the granular one
plot_motif_summary(neg_patterns %>% mutate(annotation_broad = motif_name),
                   hits_all_anno %>% mutate(annotation_broad = motif_name),
                   hit_dyad_dist %>% mutate(annotation_broad = motif_name),
                   plot_known_motifs = TRUE,
                   rel_widths = c(4, 2, 0.5, 1, 1, 1, 1, 1, 1, 1, 1),
                   to_revcomp = c("456|ZEB/SNAI", "455|YY1/2repressive"))

```



## Unresolved motifs

```{r motifs_unresolved, fig.width = 21, fig.height = 11}

unresolved <- motifs_compiled_unique %>% 
  filter(category == "unresolved") %>% 
  arrange(desc(total_hits))

plot_motif_summary(unresolved %>% mutate(annotation_broad = motif_name),
                   hits_all_anno %>% mutate(annotation_broad = motif_name), 
                   hit_dyad_dist %>% mutate(annotation_broad = motif_name),
                   order_by = "distToTSS",
                   plot_known_motifs = FALSE,
                   rel_widths = c(3.5, 0.5, 1, 1, 1, 1, 1, 1, 1, 1),
                   top_n = 30)

```



# Footprinting motif instances

For a few motifs, let's compute the ATAC footprint, focusing on the the cluster with
the most hits for each of the motifs of interest:

```{r}

# for each motif of interest, what is the cluster w/ the most hits?
hits_all_anno %>%
  group_by(annotation_broad) %>%
  slice_max(order_by = n, n = 1) %>%
  filter(annotation_broad %in%
           c("CTCF", "GATA", "BZIP", "NFY", "NFYrep", "YY1/2", "YY1/2rep", "HIC", "ZEB/SNAI", "BCL11A/IRF/SPI", "BCL11Arep")) %>%
  dplyr::select(pattern_class, motif_name, Cluster, n)

```

## Load BPCells object

```{r load_bpcells}

global_bp_obj <- readRDS(here("output/05-misc/01/global_bp_obj.rds"))

head(global_bp_obj$cell_metadata)

```


## Helper function

The motif footprint calculation counts insertions (i.e. fragment starts, ends)
that fall within each position in the window. Then the fragment count at each position
is normalized by (divided by) the mean count in the outer 10% of the window, taken as a measure of the local
background accessibility.

```{r fn_footprint}

footprint_fn <- function(cluster,
                         motif_broad,
                         split_by_quantile = TRUE,
                         sample_instances = TRUE,
                         sample_n = 10000,
                         flank = 250) {
  
  motif_safe <- gsub("\\/", ".", motif_broad)
  q_string <- ifelse(split_by_quantile, "quantiles", "noquantiles")
  s_string <- ifelse(sample_instances, glue("sample{sample_n}"), "nosample")
  out_path <- glue("{out}/{cluster}__{motif_safe}.footprint.{q_string}.{s_string}.flank{flank}.Rds")
  
  if (file.exists(out_path)) {
    
    message("@ found previous output at ", basename(out_path))
    fp_out <- readRDS(out_path)
    return(fp_out)
    
  }
  
  message("@ loading hits for cluster ", cluster)
  # load hits for the cluster of interest
  motif_instances <- rtracklayer::import.bed(glue("{chrombpnet_dir}/02-compendium/hits_unified_motifs/reconciled_per_celltype_peaks/{cluster}/{finemo_param}/{cluster}__hits_unified.{finemo_param}.reconciled.bed.gz"),
                                             extraCols = c("pattern_class" = "character"))
  
  # map in the broad annotation
  motif_instances$annotation_broad <- plyr::mapvalues(motif_instances$name,
                                                      from = motifs_compiled_unique$motif_name,
                                                      to = motifs_compiled_unique$annotation_broad,
                                                      warn_missing = FALSE)
  
  instances_motif <- motif_instances[motif_instances$annotation_broad == motif_broad]
  
  # subset the global object
  sub_meta <- global_bp_obj$cell_metadata %>%
    dplyr::filter(Cluster_chrombpnet == cluster)
  
  frags <- global_bp_obj$frags %>% select_cells(sub_meta$cb)
  
  # if there are different variants of the motif, subset to the most common one -
  # motif instances may differ because they are identified with slightly different
  # input motifs, and then reconciled afterwards to a unique compendium motif
  message(glue("@ found {length(motif_instances)} motif instances"))
  n_variants  <- instances_motif$name %>% unique %>% length
  most_common <- instances_motif %>% width() %>% table() %>% which.max() %>% names()
  instances_subset <- instances_motif[width(instances_motif) == most_common]
  message(glue("@ motif instances remaining: N={length(instances_subset)}"))
  
  # split motifs by motif score quantile
  if (split_by_quantile) {
    
    message("@ splitting hits by quantile")
    quantiles <- quantile(instances_subset$score)
    instances_subset$quantile <- cut(instances_subset$score, breaks = quantiles,
                                     include.lowest = TRUE, labels = 1:4) %>% 
      as.character()
    
    print(unique(instances_subset$quantile))
    
  } else { instances_subset$quantile <- 1 }
  
  n_quantiles <- length(unique(instances_subset$quantile))
  
  # sample motif instances
  if (sample_instances) {
    
    instances_per_q_sampled <- map(1:n_quantiles, function(q) {
      
      instances_per_q <- instances_subset[instances_subset$quantile == q]
      
      if (length(instances_per_q) < sample_n) {
        message(glue("@@ quantile {q}: <{sample_n} motif instances"))
        return(instances_per_q)
      }
      else {
        message(glue("@@ quantile {q}: sampling {sample_n} motif instances"))
        instances_per_q %>% sample(sample_n)
      }
      
    })
    
  } else {
    
    instances_per_q_sampled <- split(instances_subset, instances_subset$quantile)
    
  }
  
  print(length(instances_per_q_sampled))
  
  n_instances <- map_dbl(instances_per_q_sampled, length) %>% sum
  
  # compute footprints
  message("@ computing footprints...")
  footprints_data <- imap_dfr(instances_per_q_sampled, function(i, q) {
    
    if (split_by_quantile) message(glue("@@ quantile {q}"))
    fp <- BPCells::plot_tf_footprint(
      fragments = frags,
      motif_positions = i,
      cell_groups = sub_meta$Cluster_chrombpnet,
      flank = flank,
      smooth = 2,
      return_data = TRUE
    ) %>% 
      mutate(quantile = q)
    
    return(fp)
    
  }
  )
  
  fp_out <- list(
    "cluster"     = cluster,
    "motif"       = motif_broad,
    "footprint"   = footprints_data,
    "width"       = as.numeric(most_common),
    "n_instances" = n_instances)
  
  message("@ saving results to ", basename(out_path))
  
  saveRDS(fp_out, file = out_path)
  
  return(fp_out)
  
  message("@ done.")
  
}

plot_footprint_fn <- function(footprints_out, central_window = 500) {
  
  n_quantiles <- length(unique(footprints_out$footprint$quantile))
  
  message("@ plotting...")
  footprints_out$footprint$quantile <- as.character(footprints_out$footprint$quantile)
  
  if (max(footprints_out$footprint$quantile) > 1) blues <- RColorBrewer::brewer.pal(8, "Blues") %>%
    tail(4)
  else blues <- "black"
  
  flank <- central_window / 2
  
  # plot footprints 
  p1 <- footprints_out$footprint %>% 
    filter(position > -flank & position < flank) %>% 
    ggplot(aes(position, smoothed_enrichment, color = quantile)) +
    # highlight the motif position within the window
    annotate("rect",
             xmin = - footprints_out$width / 2, xmax = footprints_out$width / 2,
             ymin = -Inf, ymax = Inf,
             fill = "yellow", alpha = 0.3) +
    geom_line(alpha = 0.8) +
    scale_color_manual(values = blues) +
    labs(x = "Distance from motif center (bp)", y = "Enrichment") +
    ggtitle(glue("{footprints_out$motif} \n(N={footprints_out$n_instances} instances, {footprints_out$cluster})")) +
    square()
  
  if (n_quantiles == 1) p1 <- p1 + no_legend()
  
  p1
  
}

```


## CTCF

We want to get footprints of a specific TF, so we filter to instances of e.g. CTCF here,
which is a positive control that we know should footprint well in a high-quality
ATACseq dataset.

```{r ctcf_footprint, fig.width = 5, fig.height = 5}

fp_ctcf_out <- footprint_fn(cluster = "Liver_c1",
                            motif_broad = "CTCF",
                            split_by_quantile = TRUE,
                            sample_instances = FALSE)

plot_footprint_fn(fp_ctcf_out)

# no splitting
fp_ctcf_out2 <- footprint_fn(cluster = "Liver_c1",
                             motif_broad = "CTCF",
                             split_by_quantile = FALSE,
                             sample_instances = FALSE)

plot_footprint_fn(fp_ctcf_out2)

```





## NFY

```{r nfy_footprint, fig.width = 5, fig.height = 5}

fp_nfy_out1 <- footprint_fn(cluster = "Brain_c9",
                            motif_broad = "NFY",
                            split_by_quantile = TRUE,
                            sample_instances = FALSE)

plot_footprint_fn(fp_nfy_out1)

fp_nfy_out <- footprint_fn(
  cluster = "Brain_c9",
  motif_broad = "NFY",
  split_by_quantile = FALSE,
  sample_instances = FALSE
)

plot_footprint_fn(fp_nfy_out)

```



## NFY repressive


```{r nfyrep_footprint, fig.width = 5, fig.height = 5}

fp_nfyrep_out <- footprint_fn(
  cluster = "Heart_c4",
  motif_broad = "NFYrep",
  split_by_quantile = FALSE,
  sample_instances = FALSE,
  flank = 500
)

plot_footprint_fn(fp_nfyrep_out)

fp_nfyrep_out2 <- footprint_fn(
  cluster = "Heart_c4",
  motif_broad = "NFYrep",
  split_by_quantile = TRUE,
  sample_instances = FALSE,
  flank = 500
)

plot_footprint_fn(fp_nfyrep_out2)

```

## YY1/2

```{r yy1_footprint, fig.width = 5, fig.height = 5}

fp_yy1_out <- footprint_fn(
  cluster = "Lung_c1",
  motif_broad = "YY1/2",
  split_by_quantile = FALSE,
  sample_instances = FALSE,
  flank = 500
)

plot_footprint_fn(fp_yy1_out)

fp_yy1_out2 <- footprint_fn(
  cluster = "Lung_c1",
  motif_broad = "YY1/2",
  split_by_quantile = TRUE,
  sample_instances = FALSE,
  flank = 500
)

plot_footprint_fn(fp_yy1_out2)

```

## YY1/2 repressive

```{r yy1rep_footprint, fig.width = 5, fig.height = 5}

fp_yy1rep_out <- footprint_fn(
  cluster = "Liver_c1",
  motif_broad = "YY1/2rep",
  split_by_quantile = FALSE,
  sample_instances = FALSE,
  flank = 500
)

plot_footprint_fn(fp_yy1rep_out)
plot_footprint_fn(fp_yy1rep_out)

fp_yy1rep_out2 <- footprint_fn(
  cluster = "Liver_c1",
  motif_broad = "YY1/2rep",
  split_by_quantile = TRUE,
  sample_instances = FALSE,
  flank = 500
)

plot_footprint_fn(fp_yy1rep_out2)

```

## ZEB/SNAI

```{r zeb_footprint, fig.width = 5, fig.height = 5}

fp_zeb_out <- footprint_fn(
  cluster = "Skin_c1",
  motif_broad = "ZEB/SNAI",
  split_by_quantile = FALSE,
  sample_instances = FALSE,
  flank = 500
)

plot_footprint_fn(fp_zeb_out)

# split quantiles
fp_zeb_out2 <- footprint_fn(
  cluster = "Skin_c1",
  motif_broad = "ZEB/SNAI",
  split_by_quantile = TRUE,
  sample_instances = FALSE,
  flank = 500
)

plot_footprint_fn(fp_zeb_out2)

``` 

## HIC

```{r hic_footprint, fig.width = 5, fig.height = 5}

fp_hic_out <- footprint_fn(
  cluster = "Heart_c3",
  motif_broad = "HIC",
  split_by_quantile = FALSE,
  sample_instances = FALSE
)

plot_footprint_fn(fp_hic_out)

fp_hic_out2 <- footprint_fn(
  cluster = "Heart_c3",
  motif_broad = "HIC",
  split_by_quantile = TRUE,
  sample_instances = FALSE,
  flank = 500
)

plot_footprint_fn(fp_hic_out2)

```

## BCL11A

```{r bcl11a_footprint, fig.width = 5, fig.height = 5}

fp_bcl11_out <- footprint_fn(
  cluster = "Liver_c7",
  motif_broad = "BCL11A/IRF/SPI",
  split_by_quantile = FALSE,
  sample_instances = FALSE)

plot_footprint_fn(fp_bcl11_out)

fp_bcl11_out2 <- footprint_fn(
  cluster = "Liver_c7",
  motif_broad = "BCL11A/IRF/SPI",
  split_by_quantile = TRUE,
  sample_instances = FALSE,
  flank = 500
)

plot_footprint_fn(fp_bcl11_out2)

```

## BCL11A repressive

```{r bcl11arep_footprint, fig.width = 5, fig.height = 5}

fp_bcl11rep_out <- footprint_fn(
  cluster = "Skin_c2",
  motif_broad = "BCL11Arep",
  split_by_quantile = FALSE,
  sample_instances = FALSE)

plot_footprint_fn(fp_bcl11rep_out)

fp_bcl11rep_out2 <- footprint_fn(
  cluster = "Skin_c2",
  motif_broad = "BCL11Arep",
  split_by_quantile = TRUE,
  sample_instances = FALSE)

plot_footprint_fn(fp_bcl11rep_out2)

```



# Session info

```{r sinfo, cache = FALSE}

sessionInfo()

```

